---
output:  
    word_document :
        reference_docx: QaTableStyle6.docx
        toc: yes
params: 
    hospital: 
        label: "Hospital from which data is being pulled"
        value: CHCO
        input: select
        choices: [CHCO, DH]
    DBServerName: 
        label: "Database Server Name:"
        value: cc-s-d05.ucdenver.pvt
        input: select
        choices: [cc-s-d05.ucdenver.pvt, vwphbisql06, ihrsql1.ihr.or.kp.org, PRDVDWSQL01, SQL01, BI01, vwedwtstsql04]
    DBName:
        label: "Database Name:"
        value: VDW_3_1_DH
        input: select
        choices: [VDW_3_1_DH, VDW.dbo, VDW, CHORDS_VDW_CC_3.1, CHORDS_VDW_CT_3.1, CHORDS_VDW_HP_3.1, CHORDS_VDW_MC_3.1, CHORDS_VDW_SC_3.1, CHORDS_VDW_SS_3.1, CHORDSlive.dbo, VDW.aligned, CHORDSlive]
---

```{r setup, include=FALSE}
options(scipen=999, digits = 5, width = 88)
options(qwraps2_markup = "markdown")

options(stringAsFactors = F)
options(dplyr.print_max = 10)
options(dplyr.pring_min = 5)

knitr::opts_chunk$set(#cache   = TRUE,
               echo    = FALSE,
               results = "hide",
               message = FALSE,
               warning = FALSE) 
startTime <- Sys.time()

library(knitr)
library(tidyverse)

# connection to sql server
con <- DBI::dbConnect(odbc::odbc(),
                      Driver = "SQL Server",
                      Server = params$DBServerName,
                      Database = params$DBName)

con <- DBI::dbConnect(odbc::odbc(),
                      Driver   = "SQL Server",
                      Server   = "cc-s-d05.ucdenver.pvt",
                      Database = "VDW_3_1_DH_0418")

hospital <- ifelse(params$hospital == "CHCO", "Children's Hospital Colorado", "Denver Health")

hospital <- "DH"

# codes <- readr::read_csv("Q:/MattCM/CHORDS/surg2rec/inclusion codes.csv")
# devtools::use_data(codes, overwrite = T)

```

#Opioid Prescription Refills Data Tables

The data below represent data pulled for the R18 grant application for Lisa McLeod.  The data in this report come from CHORDS, specifically from `r hospital`.  The server name from which the data are pulled is `r params$DBServerName` and the database is `r params$DBName`.  

###Information about the data pull program 
Hospital: `r hospital`
Server Name: `r params$DBServerName`
Database Name: `r params$DBName`
Query Run Date:  `r Sys.Date()`

```{r supporting_data_and_functions, include = FALSE}

start_dt <- "2011-01-01"
end_dt   <- "2016-12-31"

## helper functions
 # returns a vector with the column numbers of character variables in the data frame
charVars <- function(df) grep('^ch',sapply(df,class))

 # trims all character variables in a dataframe, sets blank to NA
trimChrVars <- function(df){
for(i in charVars(df)){
  df[,i] <- gsub('\\s+$','',df[,i])
  df[,i] <- ifelse(nchar(df[,i])==0,NA,df[,i])
}
  df
}

# age category calculator
ageCatCalc <- function(age){
  ageCat <- factor(
    ifelse(is.na(age)==TRUE, 0,
           ifelse(age<0              , 1, 
                  ifelse(age>=0 & age<2, 2,
                         ifelse(age>=2 & age<5, 3, 
                                ifelse(age>=5 & age<10, 4, 
                                       ifelse (age>=10 & age<15, 5, 
                                               ifelse(age>=15 & age<19, 6,
                                                      ifelse(age>=19 & age<22, 7,
                                                             ifelse(age>=22 & age<45, 8, 
                                                                    ifelse(age>=45 & age<65, 9, 
                                                                           ifelse(age>=65 & age<75, 10, 
                                                                                  ifelse(age>=75 & age<90, 11,
                                                                                         ifelse(age>=90 , 12, 13)))))) ))))))),
    levels=0:13,
    labels=c('Missing','Negative','0-1','2-4','5-9','10-14','15-18','19-21','22-44','45-64','65-74','75-89','90+','Other')
  )
  
  return(ageCat)
}

# supporting look up tables

stateCnty <- surg2rec::stateCnty

ccsMDxI9 <- surg2rec::ccsMDxI9

ccsMDxI10 <- surg2rec::ccsMDxI10

ccsSDxI9 <- surg2rec::ccsSDxI9

icdLU <- rbind(ccsMDxI10[c('dxCdTp','dxDesc')],ccsSDxI9[c('dxCdTp','dxDesc')]) 

# ISO language lookup
isoLang <- surg2rec::isoLang 
 
# VDW value sets
valSets <- surg2rec::valSets

#Inclusion codes
codes <- surg2rec::codes

# Provider specialties
specialties <- surg2rec::specialties

# Provider type
prov_type <- surg2rec::prov_type

#Create long version of codes file
codes2 <- codes %>% 
  tidyr::gather(.) %>% 
  dplyr::filter(complete.cases(.)) %>% 
  dplyr::rename(code_type = key, code = value) %>% 
  dplyr::mutate(
    spine = if_else(stringr::str_detect(code_type, "spine"), 1, 0),
    pectus = if_else(stringr::str_detect(code_type, "pectus"), 1, 0),
    ProcedureType = if_else(spine == 1, "Spine", "Pectus")
  )


```

# Table Set 1: Facility Demographics

##Table 1A. Surgeries by Year and Surgery Type

This table contains the number of procedures performed by year and type of procedure

```{r tab1_sum_stats}
          
tab1a <- odbc::dbGetQuery(con,
       "SELECT
        DISTINCT z.ENC_ID,
        z.PERSON_ID,
       MAX(CASE
               WHEN z.PX IN('0RG4070', '0RG4071', '0RG407J', '0RG40A0', '0RG40A1', '0RG40AJ', '0RG40J0', '0RG40J1', '0RG40JJ', '0RG40K0', '0RG40K1', '0RG40KJ', '0RG40Z0', '0RG40Z1', '0RG40ZJ', '0RG6070', '0RG6071', '0RG607J', '0RG60A0', '0RG60A1', '0RG60AJ', '0RG60J0', '0RG60J1', '0RG60JJ', '0RG60K0', '0RG60K1', '0RG60KJ', '0RG60Z0', '0RG60Z1', '0RG60ZJ', '0RG7070', '0RG7071', '0RG707J', '0RG70A0', '0RG70A1', '0RG70AJ', '0RG70J0', '0RG70J1', '0RG70JJ', '0RG70K0', '0RG70K1', '0RG70KJ', '0RG70Z0', '0RG70Z1', '0RG70ZJ', '0RG8070', '0RG8071', '0RG807J', '0RG80A0', '0RG80A1', '0RG80AJ', '0RG80J0', '0RG80J1', '0RG80JJ', '0RG80K0', '0RG80K1', '0RG80KJ', '0RG80Z0', '0RG80Z1', '0RG80ZJ', '0RGA070', '0RGA071', '0RGA07J', '0RGA0A0', '0RGA0A1', '0RGA0AJ', '0RGA0J0', '0RGA0J1', '0RGA0JJ', '0RGA0K0', '0RGA0K1', '0RGA0KJ', '0RGA0Z0', '0RGA0Z1', '0RGA0ZJ', '0SG0070', '0SG0071', '0SG007J', '0SG00A0', '0SG00A1', '0SG00AJ', '0SG00J0', '0SG00J1', '0SG00JJ', '0SG00K0', '0SG00K1', '0SG00KJ', '0SG00Z0', '0SG00Z1', '0SG00ZJ', '0SG1070', '0SG1071', '0SG107J', '0SG10A0', '0SG10A1', '0SG10AJ', '0SG10J0', '0SG10J1', '0SG10JJ', '0SG10K0', '0SG10K1', '0SG10KJ', '0SG10Z0', '0SG10Z1', '0SG10ZJ', '0SG3070', '0SG3071', '0SG307J', '0SG30A0', '0SG30A1', '0SG30AJ', '0SG30J0', '0SG30J1', '0SG30JJ', '0SG30K0', '0SG30K1', '0SG30KJ', '0SG30Z0', '0SG30Z1', '0SG30ZJ', '81', '81.01', '81.02', '81.03', '81.04', '81.05', '81.06', '81.07', '81.08', '81.09', '81.3', '81.35', '81.36', '81.37', '81.38', '81.39', '81.61', '81.62', '81.63', '81.64', '84.51', '22532', '22533', '22548', '22551', '22554', '22556', '22558', '22586', '22590', '22595', '22600', '22610', '22612', '22614', '22630', '22633', '22634', '22800', '22802', '22804', '22808', '22810', '22812', '22818', '22819', '22830', '22840', '22841', '22842', '22843', '22844', '22845', '22846', '22847', '22848', '22849', '22850', '22852', '22853', '22854', '22855', '22856', '22857', '22858', '22859', '22861', '22862', '22864', '22865', '22867', '22868', '22869', '22870')
               THEN 1 ELSE 0 END) AS spine,
      MAX(CASE
               WHEN z.PX IN('0WU80JZ', '0WU84JZ', '34.74', '21740', '21742', '21743')
               THEN 1 ELSE 0 END) AS pectus,
      z.PROCDATE,
      YEAR(z.ADATE) as px_date
FROM
         DEMOGRAPHICS a
         INNER JOIN [PROCEDURES] z
              ON z.PERSON_ID = a.PERSON_ID
    WHERE  z.ADATE >= '01/01/2011'
           AND DATEDIFF(YY, a.BIRTH_DATE, z.ADATE) - CASE
                      WHEN DATEADD(YY, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE), a.BIRTH_DATE) > z.ADATE
                      THEN 1
                      ELSE 0
                    END >= 10
           AND DATEDIFF(YY, a.BIRTH_DATE, z.ADATE) - CASE
                      WHEN DATEADD(YY, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE), a.BIRTH_DATE) > z.ADATE
                      THEN 1
                      ELSE 0
                    END <= 26
           AND 
               z.PX IN('0RG4070', '0RG4071', '0RG407J', '0RG40A0', '0RG40A1', '0RG40AJ', '0RG40J0', '0RG40J1', '0RG40JJ', '0RG40K0', '0RG40K1', '0RG40KJ', '0RG40Z0', '0RG40Z1', '0RG40ZJ', '0RG6070', '0RG6071', '0RG607J', '0RG60A0', '0RG60A1', '0RG60AJ', '0RG60J0', '0RG60J1', '0RG60JJ', '0RG60K0', '0RG60K1', '0RG60KJ', '0RG60Z0', '0RG60Z1', '0RG60ZJ', '0RG7070', '0RG7071', '0RG707J', '0RG70A0', '0RG70A1', '0RG70AJ', '0RG70J0', '0RG70J1', '0RG70JJ', '0RG70K0', '0RG70K1', '0RG70KJ', '0RG70Z0', '0RG70Z1', '0RG70ZJ', '0RG8070', '0RG8071', '0RG807J', '0RG80A0', '0RG80A1', '0RG80AJ', '0RG80J0', '0RG80J1', '0RG80JJ', '0RG80K0', '0RG80K1', '0RG80KJ', '0RG80Z0', '0RG80Z1', '0RG80ZJ', '0RGA070', '0RGA071', '0RGA07J', '0RGA0A0', '0RGA0A1', '0RGA0AJ', '0RGA0J0', '0RGA0J1', '0RGA0JJ', '0RGA0K0', '0RGA0K1', '0RGA0KJ', '0RGA0Z0', '0RGA0Z1', '0RGA0ZJ', '0SG0070', '0SG0071', '0SG007J', '0SG00A0', '0SG00A1', '0SG00AJ', '0SG00J0', '0SG00J1', '0SG00JJ', '0SG00K0', '0SG00K1', '0SG00KJ', '0SG00Z0', '0SG00Z1', '0SG00ZJ', '0SG1070', '0SG1071', '0SG107J', '0SG10A0', '0SG10A1', '0SG10AJ', '0SG10J0', '0SG10J1', '0SG10JJ', '0SG10K0', '0SG10K1', '0SG10KJ', '0SG10Z0', '0SG10Z1', '0SG10ZJ', '0SG3070', '0SG3071', '0SG307J', '0SG30A0', '0SG30A1', '0SG30AJ', '0SG30J0', '0SG30J1', '0SG30JJ', '0SG30K0', '0SG30K1', '0SG30KJ', '0SG30Z0', '0SG30Z1', '0SG30ZJ', '81', '81.01', '81.02', '81.03', '81.04', '81.05', '81.06', '81.07', '81.08', '81.09', '81.3', '81.35', '81.36', '81.37', '81.38', '81.39', '81.61', '81.62', '81.63', '81.64', '84.51', '0WU80JZ', '0WU84JZ', '34.74', '22532', '22533', '22548', '22551', '22554', '22556', '22558', '22586', '22590', '22595', '22600', '22610', '22612', '22614', '22630', '22633', '22634', '22800', '22802', '22804', '22808', '22810', '22812', '22818', '22819', '22830', '22840', '22841', '22842', '22843', '22844', '22845', '22846', '22847', '22848', '22849', '22850', '22852', '22853', '22854', '22855', '22856', '22857', '22858', '22859', '22861', '22862', '22864', '22865', '22867', '22868', '22869', '22870', '21740', '21742', '21743')
GROUP BY z.PERSON_ID, z.ENC_ID, z.PROCDATE, YEAR(z.ADATE)"

)

tab1a2 <- tab1a %>% 
  group_by(PERSON_ID) %>% 
  mutate(
    min_dt = min(PROCDATE),
    use_dt = if_else(min_dt == PROCDATE, 1, 0)
  ) %>% 
  ungroup(.) %>% 
  dplyr::filter(use_dt == 1) %>% 
  group_by(px_date) %>% 
  summarise(spine = sum(spine), pectus = sum(pectus)) %>% 
  ungroup(.)

tab1a_tot <- odbc::dbGetQuery(con,
       "SELECT
      count(*) as n_pxs,
      YEAR(z.ADATE) as px_date
    FROM
         DEMOGRAPHICS a
         JOIN [PROCEDURES] z
              ON z.PERSON_ID = a.PERSON_ID
    WHERE  z.ADATE >= '01/01/2011'
           AND DATEDIFF(YY, a.BIRTH_DATE, z.ADATE) - CASE
                      WHEN DATEADD(YY, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE), a.BIRTH_DATE) > z.ADATE
                      THEN 1
                      ELSE 0
                    END >= 10
           AND DATEDIFF(YY, a.BIRTH_DATE, z.ADATE) - CASE
                      WHEN DATEADD(YY, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE), a.BIRTH_DATE) > z.ADATE
                      THEN 1
                      ELSE 0
                    END <= 26 
GROUP BY YEAR(z.ADATE)
ORDER BY YEAR(z.ADATE)"
)

tab1a_all <- tab1a2 %>% 
  left_join(tab1a_tot, by = "px_date") %>% 
  mutate(
    spine2  = paste0(spine, " (", round((spine/n_pxs) * 100, 2), "%)"),
    pectus2 = paste0(pectus, " (", round((pectus/n_pxs) * 100, 2), "%)"),
    total   = paste0(spine + pectus, " (", round(((spine + pectus)/n_pxs) * 100, 2), "%)")
  ) %>% 
  select(px_date, spine2, pectus2, total)

```

```{r tab1a_out, results="asis"}

knitr::kable(tab1a_all, col.names = c("Year", "Spine", "Pectus", "Total"))

```

##Table 1B: Cumulative Demographic Demographic Differences by Facility

Heights and weights in this table represent an average of the heights and weights recorded during the procedure encounter.  This was done because 1) height and weight weren't measured at the same time for some individuals and 2) height and weight weren't measured during consistent times for each participant.  

```{r tab1b}

proc_test <- odbc::dbGetQuery(con,
                                 "SELECT DISTINCT(PROCEDURES_ID)
                                 FROM PROCEDURES")

proc_test2 <- odbc::dbGetQuery(con,
                                 "SELECT DISTINCT(ENC_ID)
                                 FROM PROCEDURES")
                                 

tab1b <- odbc::dbGetQuery(con,
       "SELECT
        DISTINCT z.ENC_ID,
        z.PERSON_ID,
       MAX(CASE
               WHEN z.PX IN('0RG4070', '0RG4071', '0RG407J', '0RG40A0', '0RG40A1', '0RG40AJ', '0RG40J0', '0RG40J1', '0RG40JJ', '0RG40K0', '0RG40K1', '0RG40KJ', '0RG40Z0', '0RG40Z1', '0RG40ZJ', '0RG6070', '0RG6071', '0RG607J', '0RG60A0', '0RG60A1', '0RG60AJ', '0RG60J0', '0RG60J1', '0RG60JJ', '0RG60K0', '0RG60K1', '0RG60KJ', '0RG60Z0', '0RG60Z1', '0RG60ZJ', '0RG7070', '0RG7071', '0RG707J', '0RG70A0', '0RG70A1', '0RG70AJ', '0RG70J0', '0RG70J1', '0RG70JJ', '0RG70K0', '0RG70K1', '0RG70KJ', '0RG70Z0', '0RG70Z1', '0RG70ZJ', '0RG8070', '0RG8071', '0RG807J', '0RG80A0', '0RG80A1', '0RG80AJ', '0RG80J0', '0RG80J1', '0RG80JJ', '0RG80K0', '0RG80K1', '0RG80KJ', '0RG80Z0', '0RG80Z1', '0RG80ZJ', '0RGA070', '0RGA071', '0RGA07J', '0RGA0A0', '0RGA0A1', '0RGA0AJ', '0RGA0J0', '0RGA0J1', '0RGA0JJ', '0RGA0K0', '0RGA0K1', '0RGA0KJ', '0RGA0Z0', '0RGA0Z1', '0RGA0ZJ', '0SG0070', '0SG0071', '0SG007J', '0SG00A0', '0SG00A1', '0SG00AJ', '0SG00J0', '0SG00J1', '0SG00JJ', '0SG00K0', '0SG00K1', '0SG00KJ', '0SG00Z0', '0SG00Z1', '0SG00ZJ', '0SG1070', '0SG1071', '0SG107J', '0SG10A0', '0SG10A1', '0SG10AJ', '0SG10J0', '0SG10J1', '0SG10JJ', '0SG10K0', '0SG10K1', '0SG10KJ', '0SG10Z0', '0SG10Z1', '0SG10ZJ', '0SG3070', '0SG3071', '0SG307J', '0SG30A0', '0SG30A1', '0SG30AJ', '0SG30J0', '0SG30J1', '0SG30JJ', '0SG30K0', '0SG30K1', '0SG30KJ', '0SG30Z0', '0SG30Z1', '0SG30ZJ', '81', '81.01', '81.02', '81.03', '81.04', '81.05', '81.06', '81.07', '81.08', '81.09', '81.3', '81.35', '81.36', '81.37', '81.38', '81.39', '81.61', '81.62', '81.63', '81.64', '84.51', '22532', '22533', '22548', '22551', '22554', '22556', '22558', '22586', '22590', '22595', '22600', '22610', '22612', '22614', '22630', '22633', '22634', '22800', '22802', '22804', '22808', '22810', '22812', '22818', '22819', '22830', '22840', '22841', '22842', '22843', '22844', '22845', '22846', '22847', '22848', '22849', '22850', '22852', '22853', '22854', '22855', '22856', '22857', '22858', '22859', '22861', '22862', '22864', '22865', '22867', '22868', '22869', '22870')
               THEN 1 ELSE 0 END) AS spine,
      MAX(CASE
               WHEN z.PX IN('0WU80JZ', '0WU84JZ', '34.74', '21740', '21742', '21743')
               THEN 1 ELSE 0 END) AS pectus,
      DATEDIFF(YY, a.BIRTH_DATE, z.ADATE) - CASE
                      WHEN DATEADD(YY, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE), a.BIRTH_DATE) > z.ADATE
                      THEN 1
                      ELSE 0 
                  END AS [AGE],
      z.ADATE,
      z.PROCDATE,
      a.GENDER,
      a.PRIMARY_LANGUAGE,
      a.RACE1, 
      b.HT,
      b.WT,
      b.MEASURE_DATE,
      c.GEOCODE,
      c.LOC_START,
      c.LOC_END,
      d.DDATE
FROM
         DEMOGRAPHICS a
         INNER JOIN [PROCEDURES] z
              ON z.PERSON_ID = a.PERSON_ID
         LEFT JOIN [ENCOUNTERS] d
              ON z.PERSON_ID = d.PERSON_ID AND z.ENC_ID = d.ENC_ID
         LEFT JOIN [VITAL_SIGNS] b
              ON z.PERSON_ID = b.PERSON_ID AND z.ENC_ID = b.ENC_ID
         LEFT JOIN [CENSUS_LOCATION] c
              ON z.PERSON_ID = c.PERSON_ID AND 
                CASE WHEN c.LOC_END IS NOT NULL AND z.ADATE BETWEEN c.LOC_START 
                  AND c.LOC_END THEN 1 WHEN c.LOC_END IS NULL AND c.LOC_START <= z.ADATE THEN 1 ELSE 0 END = 1
    WHERE  z.ADATE >= '01/01/2011'
           AND DATEDIFF(YY, a.BIRTH_DATE, z.ADATE) - CASE
                      WHEN DATEADD(YY, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE), a.BIRTH_DATE) > z.ADATE
                      THEN 1
                      ELSE 0
                    END >= 10
           AND DATEDIFF(YY, a.BIRTH_DATE, z.ADATE) - CASE
                      WHEN DATEADD(YY, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE), a.BIRTH_DATE) > z.ADATE
                      THEN 1
                      ELSE 0
                    END <= 26
           AND 
               z.PX IN('0RG4070', '0RG4071', '0RG407J', '0RG40A0', '0RG40A1', '0RG40AJ', '0RG40J0', '0RG40J1', '0RG40JJ', '0RG40K0', '0RG40K1', '0RG40KJ', '0RG40Z0', '0RG40Z1', '0RG40ZJ', '0RG6070', '0RG6071', '0RG607J', '0RG60A0', '0RG60A1', '0RG60AJ', '0RG60J0', '0RG60J1', '0RG60JJ', '0RG60K0', '0RG60K1', '0RG60KJ', '0RG60Z0', '0RG60Z1', '0RG60ZJ', '0RG7070', '0RG7071', '0RG707J', '0RG70A0', '0RG70A1', '0RG70AJ', '0RG70J0', '0RG70J1', '0RG70JJ', '0RG70K0', '0RG70K1', '0RG70KJ', '0RG70Z0', '0RG70Z1', '0RG70ZJ', '0RG8070', '0RG8071', '0RG807J', '0RG80A0', '0RG80A1', '0RG80AJ', '0RG80J0', '0RG80J1', '0RG80JJ', '0RG80K0', '0RG80K1', '0RG80KJ', '0RG80Z0', '0RG80Z1', '0RG80ZJ', '0RGA070', '0RGA071', '0RGA07J', '0RGA0A0', '0RGA0A1', '0RGA0AJ', '0RGA0J0', '0RGA0J1', '0RGA0JJ', '0RGA0K0', '0RGA0K1', '0RGA0KJ', '0RGA0Z0', '0RGA0Z1', '0RGA0ZJ', '0SG0070', '0SG0071', '0SG007J', '0SG00A0', '0SG00A1', '0SG00AJ', '0SG00J0', '0SG00J1', '0SG00JJ', '0SG00K0', '0SG00K1', '0SG00KJ', '0SG00Z0', '0SG00Z1', '0SG00ZJ', '0SG1070', '0SG1071', '0SG107J', '0SG10A0', '0SG10A1', '0SG10AJ', '0SG10J0', '0SG10J1', '0SG10JJ', '0SG10K0', '0SG10K1', '0SG10KJ', '0SG10Z0', '0SG10Z1', '0SG10ZJ', '0SG3070', '0SG3071', '0SG307J', '0SG30A0', '0SG30A1', '0SG30AJ', '0SG30J0', '0SG30J1', '0SG30JJ', '0SG30K0', '0SG30K1', '0SG30KJ', '0SG30Z0', '0SG30Z1', '0SG30ZJ', '81', '81.01', '81.02', '81.03', '81.04', '81.05', '81.06', '81.07', '81.08', '81.09', '81.3', '81.35', '81.36', '81.37', '81.38', '81.39', '81.61', '81.62', '81.63', '81.64', '84.51', '0WU80JZ', '0WU84JZ', '34.74', '22532', '22533', '22548', '22551', '22554', '22556', '22558', '22586', '22590', '22595', '22600', '22610', '22612', '22614', '22630', '22633', '22634', '22800', '22802', '22804', '22808', '22810', '22812', '22818', '22819', '22830', '22840', '22841', '22842', '22843', '22844', '22845', '22846', '22847', '22848', '22849', '22850', '22852', '22853', '22854', '22855', '22856', '22857', '22858', '22859', '22861', '22862', '22864', '22865', '22867', '22868', '22869', '22870', '21740', '21742', '21743')
GROUP BY z.PERSON_ID, z.ENC_ID, z.PROCDATE, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE) - CASE
                      WHEN DATEADD(YY, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE), a.BIRTH_DATE) > z.ADATE
                      THEN 1
                      ELSE 0 
                  END, z.ADATE, a.GENDER, a.PRIMARY_LANGUAGE, a.RACE1, b.HT, b.WT, b.MEASURE_DATE, c.GEOCODE, c.LOC_START, c.LOC_END, d.DDATE"
)

tab1b$stateCnty <-ifelse(tab1b$GEOCODE == "", NA, substr(tab1b$GEOCODE, 1, 5))

## Change code below when package is created
states <- chordsTables::stateCnty #%>% 
  #select(state, stateFP, countyFP, stateCnty) 

tab1b2 <- tab1b %>% 
  left_join(., states, by = "stateCnty") %>%   
  group_by(PERSON_ID) %>%  
  mutate( 
    end_geodt = if_else(is.na(LOC_END), as.Date("2018-05-09"), as.Date(LOC_END)), 
    geo_adate = if_else(ADATE <= end_geodt, 1, 0),
    date_diff = as.Date(PROCDATE) - as.Date(MEASURE_DATE),
    miss_htwt = if_else(!is.na(HT) & !is.na(WT), 1, 0),
    min_diff  = min(abs(date_diff), na.rm = T),
    use_date  = if_else(as.Date(PROCDATE) + min_diff == MEASURE_DATE, 1, 0),
    dup_id    = if_else(duplicated(ENC_ID), 1, 0),
    avg_ht    = mean(HT, na.rm = T),
    avg_wt    = mean(WT, na.rm = T),
    miss_avg  = if_else(!is.na(avg_ht) & !is.na(avg_wt), 1, 0),
    min_yr    = min(ADATE),
    use_yr    = if_else(min_yr == ADATE, 1, 0),
    ProcedureType = if_else(spine == 1, "Spine", "Pectus")
  ) %>%  
  ungroup(.) %>% 
  dplyr::filter(use_yr == 1) %>% 
  dplyr::filter(!duplicated(PERSON_ID) & !duplicated(ENC_ID))

tab1b_sum <- 
    list("Age" =
       list("Mean (SD)" = ~ qwraps2::mean_sd(AGE, na_rm = T),
            "Median (IQR)"    = ~ qwraps2::median_iqr(AGE, na_rm = T)),
       "Gender, Freqency(%)" =
       list("Male" = ~ qwraps2::n_perc(GENDER == "M", na_rm = T)),
       "Primary Language, Freqency(%)" =
       list("English" = ~ qwraps2::n_perc(PRIMARY_LANGUAGE == "eng", na_rm = T),
            "Spanish" = ~ qwraps2::n_perc(PRIMARY_LANGUAGE == "spa", na_rm = T)),
       "Race, Freqency(%)" =
       list("Native Hawaiian or Other Pacific Islander"   = ~ qwraps2::n_perc(RACE1 == "HP", na_rm = T),
            "American Indian/Alaska Native"               = ~ qwraps2::n_perc(RACE1 == "IN", na_rm = T),
            "Asian"                                       = ~ qwraps2::n_perc(RACE1 == "AS", na_rm = T),
            "Black or African American"                   = ~ qwraps2::n_perc(RACE1 == "BA", na_rm = T),
            "White"                                       = ~ qwraps2::n_perc(RACE1 == "WH", na_rm = T),
            "More than one race"                          = ~ qwraps2::n_perc(RACE1 == "MU", na_rm = T),
            "Unknown"                                     = ~ qwraps2::n_perc(RACE1 == "UN", na_rm = T)),
       "Height" =
       list("Mean (SD)" = ~ qwraps2::mean_sd(avg_ht, na_rm = T)),
       "Weight" = 
       list("Mean (SD)" = ~ qwraps2::mean_sd(avg_wt, na_rm = T)),
       "State of patient address" = 
         list("Colorado"     = ~ qwraps2::n_perc(state == "CO", na_rm = T),
              "Wyoming"      = ~ qwraps2::n_perc(state == "WY", na_rm = T),
              "Nebraska"     = ~ qwraps2::n_perc(state == "NE", na_rm = T),
              "New Mexico"   = ~ qwraps2::n_perc(state == "NM", na_rm = T),
              "Oklahoma"     = ~ qwraps2::n_perc(state == "OK", na_rm = T),
              "South Dakota" = ~ qwraps2::n_perc(state == "SD", na_rm = T),
              "Kansas"       = ~ qwraps2::n_perc(state == "KS", na_rm = T),
              "Other/Missing"= ~ qwraps2::n_perc((!(state %in% c("CO", "WY", "NE", "NM", "OK", "SD", "KS") & !is.na(state))) , 
                                                 na_rm = T))
    )


```

```{r tab1b_disp, results="asis"}

if(length(dimnames(table(tab1b2$ProcedureType))) == 1){
  tab1b_tot_sum <- qwraps2::summary_table(tab1b2, tab1b_sum)
  tab1b_out <- tab1b_tot_sum
  print(tab1b_out, cnames = dimnames(table(tab1b2$ProcedureType))[[1]][1])
} else{
  tab1b_tot_sum <- qwraps2::summary_table(tab1b2, tab1b_sum)
  tab1b_proc_sum <- qwraps2::summary_table(dplyr::group_by(tab1b2, "ProcedureType"), tab1b_sum)
  tab1b_out <- cbind(tab1b_tot_sum, tab1b_proc_sum)
  print(tab1b_out, cnames = c("Total", dimnames(table(tab1b2$ProcedureType))[[1]]))
}


```

##Table 1C: Patient Location by Year

```{r tab1c_out, results="asis"}

tab1c <- tab1b2 %>% 
  dplyr::filter(state == "CO") %>% 
  mutate(
    YEAR = lubridate::year(PROCDATE)
  ) %>% 
  group_by(YEAR)

qwraps2::summary_table(tab1c, list("County of Residence" = with(tab1c, qwraps2::tab_summary(countyName, n_perc_args = list(digits = 2, show_symbol = TRUE))))) 

```

## Table 1D: Length of Stay by Location & Type Summary (All Years)

```{r tab1d_data}

# Some length discharge values are 1900-01-01 so a new variable needs to be created to turn values < 0 to missing for 
# length of stay

tab1d <- tab1b2 %>% 
  mutate(
    los  = as.Date(DDATE) - as.Date(ADATE),
    los2 = if_else(los < 0, NA_integer_, as.integer(los))
  )

tab1d_sum <- list(
  "Length of Stay (Days)" =
    list(
      "Mean (SD)"    =~ qwraps2::mean_sd(los2, na_rm = T),
      "Median (IQR)" =~ qwraps2::median_iqr(los2, na_rm = T),
      "Range"        =~ paste(min(los2, na.rm = T), " - ", max(los2, na.rm = T))
    )
)

```

```{r out_tab1d, results="asis"}

if(length(dimnames(table(tab1d$ProcedureType))) == 1){
  tab1d_tot_sum <- qwraps2::summary_table(tab1d, tab1d_sum)
  tab1d_out <- tab1d_tot_sum
  print(tab1d_out, cnames = dimnames(table(tab1d$ProcedureType))[[1]][1])
} else{
  tab1d_tot_sum <- qwraps2::summary_table(tab1d, tab1d_sum)
  tab1d_proc_sum <- qwraps2::summary_table(dplyr::group_by(tab1d, ProcedureType), tab1d_sum)
  tab1d_out <- cbind(tab1d_tot_sum, tab1d_proc_sum)
  print(tab1b_out, cnames = c("Total", dimnames(table(tab1b2$ProcedureType))[[1]]))
}

```

## Table Pack 1E: Length of Stay by Location and Type

```{r tab1e_data, results="asis"}

tab1e <- tab1d %>% 
  mutate(
    YEAR = lubridate::year(PROCDATE)
  )

if(length(dimnames(table(tab1e$ProcedureType))) == 1){
  tab1e_tot_sum <- qwraps2::summary_table(dplyr::group_by(tab1e, YEAR), tab1d_sum)
  tab1e_out <- tab1e_tot_sum
  print(tab1e_out)
} else{
  tab1e_tot_sum <- qwraps2::summary_table(tab1e, tab1e_sum)
  tab1e_proc_sum <- qwraps2::summary_table(dplyr::group_by(tab1e, ProcedureType, YEAR), tab1d_sum)
  tab1e_out <- cbind(tab1e_tot_sum, tab1e_proc_sum)
  print(tab1b_out, cnames = c("Total", dimnames(table(tab1b2$ProcedureType))[[1]]))
}

```

# Figure 1A: Histogram of Patient Age at surgery (CPT Code/Proc Code Date)

```{r fig1a_data}

fig1a <- ggplot(data = tab1b2, aes(x = AGE)) +
  geom_histogram(color="black", fill="red") +
  ylab("Count") + xlab("Age")

fig1a

```

#Data Tables Protocol: Table Set 2: Patient Demographic Tables

##Table Pack 2A: Diagnosis Code Frequencies by Location

```{r tab2a_data}

tab2a <- odbc::dbGetQuery(con,
       "SELECT DISTINCT z.ENC_ID,
        z.PERSON_ID,
        d.DIAGNOSES_ID,
        d.DX,
        d.PRIMARY_DX,
        d.PRINCIPAL_DX
FROM
         DEMOGRAPHICS a
         INNER JOIN [PROCEDURES] z
              ON z.PERSON_ID = a.PERSON_ID
         LEFT JOIN [DIAGNOSES] d
              ON z.PERSON_ID = d.PERSON_ID AND z.ENC_ID = d.ENC_ID
    WHERE  z.ADATE >= '01/01/2011'
           AND DATEDIFF(YY, a.BIRTH_DATE, z.ADATE) - CASE
                      WHEN DATEADD(YY, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE), a.BIRTH_DATE) > z.ADATE
                      THEN 1
                      ELSE 0
                    END >= 10
           AND DATEDIFF(YY, a.BIRTH_DATE, z.ADATE) - CASE
                      WHEN DATEADD(YY, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE), a.BIRTH_DATE) > z.ADATE
                      THEN 1
                      ELSE 0
                    END <= 26
           AND 
               z.PX IN('0RG4070', '0RG4071', '0RG407J', '0RG40A0', '0RG40A1', '0RG40AJ', '0RG40J0', '0RG40J1', '0RG40JJ', '0RG40K0', '0RG40K1', '0RG40KJ', '0RG40Z0', '0RG40Z1', '0RG40ZJ', '0RG6070', '0RG6071', '0RG607J', '0RG60A0', '0RG60A1', '0RG60AJ', '0RG60J0', '0RG60J1', '0RG60JJ', '0RG60K0', '0RG60K1', '0RG60KJ', '0RG60Z0', '0RG60Z1', '0RG60ZJ', '0RG7070', '0RG7071', '0RG707J', '0RG70A0', '0RG70A1', '0RG70AJ', '0RG70J0', '0RG70J1', '0RG70JJ', '0RG70K0', '0RG70K1', '0RG70KJ', '0RG70Z0', '0RG70Z1', '0RG70ZJ', '0RG8070', '0RG8071', '0RG807J', '0RG80A0', '0RG80A1', '0RG80AJ', '0RG80J0', '0RG80J1', '0RG80JJ', '0RG80K0', '0RG80K1', '0RG80KJ', '0RG80Z0', '0RG80Z1', '0RG80ZJ', '0RGA070', '0RGA071', '0RGA07J', '0RGA0A0', '0RGA0A1', '0RGA0AJ', '0RGA0J0', '0RGA0J1', '0RGA0JJ', '0RGA0K0', '0RGA0K1', '0RGA0KJ', '0RGA0Z0', '0RGA0Z1', '0RGA0ZJ', '0SG0070', '0SG0071', '0SG007J', '0SG00A0', '0SG00A1', '0SG00AJ', '0SG00J0', '0SG00J1', '0SG00JJ', '0SG00K0', '0SG00K1', '0SG00KJ', '0SG00Z0', '0SG00Z1', '0SG00ZJ', '0SG1070', '0SG1071', '0SG107J', '0SG10A0', '0SG10A1', '0SG10AJ', '0SG10J0', '0SG10J1', '0SG10JJ', '0SG10K0', '0SG10K1', '0SG10KJ', '0SG10Z0', '0SG10Z1', '0SG10ZJ', '0SG3070', '0SG3071', '0SG307J', '0SG30A0', '0SG30A1', '0SG30AJ', '0SG30J0', '0SG30J1', '0SG30JJ', '0SG30K0', '0SG30K1', '0SG30KJ', '0SG30Z0', '0SG30Z1', '0SG30ZJ', '81', '81.01', '81.02', '81.03', '81.04', '81.05', '81.06', '81.07', '81.08', '81.09', '81.3', '81.35', '81.36', '81.37', '81.38', '81.39', '81.61', '81.62', '81.63', '81.64', '84.51', '0WU80JZ', '0WU84JZ', '34.74', '22532', '22533', '22548', '22551', '22554', '22556', '22558', '22586', '22590', '22595', '22600', '22610', '22612', '22614', '22630', '22633', '22634', '22800', '22802', '22804', '22808', '22810', '22812', '22818', '22819', '22830', '22840', '22841', '22842', '22843', '22844', '22845', '22846', '22847', '22848', '22849', '22850', '22852', '22853', '22854', '22855', '22856', '22857', '22858', '22859', '22861', '22862', '22864', '22865', '22867', '22868', '22869', '22870', '21740', '21742', '21743')"
)

length(unique(tab2a$PERSON_ID))

los_tab <- tab1d %>% 
  select(PERSON_ID, ProcedureType, los2)

#The code in the mutate section below counts the different responses from PRIMARY_DX
#In order to determine if any secondary diagnoses can be used enlui of primary dxs
#There are no secondary diagnoses without a primary diagnosis in the DH Date

tab2a2 <- tab2a %>% 
  dplyr::filter(PRIMARY_DX == "P") %>% 
  left_join(., los_tab, by = "PERSON_ID")

tab2_test <- as.data.frame(table(tab2a2$ProcedureType, tab2a2$DX))

list_tab2 <- split(tab2_test[, c(1,2)], seq(nrow(tab2_test[, c(1,2)])))

dx_tab <- data.frame(avg_los = unlist(lapply(list_tab2, 
                                      function(x) mean(tab2a2[tab2a2$ProcedureType == x[[1]] & tab2a2$DX == x[[2]],
                                                              "los2"], na.rm = T))))

dx_tab2 <- data.frame(med_los = unlist(lapply(list_tab2, 
                                      function(x) median(tab2a2[tab2a2$ProcedureType == x[[1]] & tab2a2$DX == x[[2]],
                                                              "los2"], na.rm = T))))

tab2a3 <- bind_cols(tab2_test, dx_tab)
tab2a4 <- bind_cols(tab2a3, dx_tab2) %>% 
  dplyr::rename(ProcedureType = Var1, DX = Var2)


```

```{r tab2a_out, results="asis"}
knitr::kable(tab2a4, col.names = c("Procedure Type", "Primary DX", "Count of DX", "Mean Length of Stay", "Median Length of Stay"))

```

##Table Pack 2B: Procedure Code Frequencies & LOS by Location (TABLE NOT POPULATED)

Per email with Lisa McLeod on 05/14/2018, this table is not populated.

Need to add in count of unique combination of procedures and mean/median length of stay

```{r, tab2b_data, eval=FALSE}

tab2b <- odbc::dbGetQuery(con,
       "SELECT
        DISTINCT z.ENC_ID,
        z.PERSON_ID,
        z.PX,
       MAX(CASE
               WHEN z.PX IN('0RG4070', '0RG4071', '0RG407J', '0RG40A0', '0RG40A1', '0RG40AJ', '0RG40J0', '0RG40J1', '0RG40JJ', '0RG40K0', '0RG40K1', '0RG40KJ', '0RG40Z0', '0RG40Z1', '0RG40ZJ', '0RG6070', '0RG6071', '0RG607J', '0RG60A0', '0RG60A1', '0RG60AJ', '0RG60J0', '0RG60J1', '0RG60JJ', '0RG60K0', '0RG60K1', '0RG60KJ', '0RG60Z0', '0RG60Z1', '0RG60ZJ', '0RG7070', '0RG7071', '0RG707J', '0RG70A0', '0RG70A1', '0RG70AJ', '0RG70J0', '0RG70J1', '0RG70JJ', '0RG70K0', '0RG70K1', '0RG70KJ', '0RG70Z0', '0RG70Z1', '0RG70ZJ', '0RG8070', '0RG8071', '0RG807J', '0RG80A0', '0RG80A1', '0RG80AJ', '0RG80J0', '0RG80J1', '0RG80JJ', '0RG80K0', '0RG80K1', '0RG80KJ', '0RG80Z0', '0RG80Z1', '0RG80ZJ', '0RGA070', '0RGA071', '0RGA07J', '0RGA0A0', '0RGA0A1', '0RGA0AJ', '0RGA0J0', '0RGA0J1', '0RGA0JJ', '0RGA0K0', '0RGA0K1', '0RGA0KJ', '0RGA0Z0', '0RGA0Z1', '0RGA0ZJ', '0SG0070', '0SG0071', '0SG007J', '0SG00A0', '0SG00A1', '0SG00AJ', '0SG00J0', '0SG00J1', '0SG00JJ', '0SG00K0', '0SG00K1', '0SG00KJ', '0SG00Z0', '0SG00Z1', '0SG00ZJ', '0SG1070', '0SG1071', '0SG107J', '0SG10A0', '0SG10A1', '0SG10AJ', '0SG10J0', '0SG10J1', '0SG10JJ', '0SG10K0', '0SG10K1', '0SG10KJ', '0SG10Z0', '0SG10Z1', '0SG10ZJ', '0SG3070', '0SG3071', '0SG307J', '0SG30A0', '0SG30A1', '0SG30AJ', '0SG30J0', '0SG30J1', '0SG30JJ', '0SG30K0', '0SG30K1', '0SG30KJ', '0SG30Z0', '0SG30Z1', '0SG30ZJ', '81', '81.01', '81.02', '81.03', '81.04', '81.05', '81.06', '81.07', '81.08', '81.09', '81.3', '81.35', '81.36', '81.37', '81.38', '81.39', '81.61', '81.62', '81.63', '81.64', '84.51', '22532', '22533', '22548', '22551', '22554', '22556', '22558', '22586', '22590', '22595', '22600', '22610', '22612', '22614', '22630', '22633', '22634', '22800', '22802', '22804', '22808', '22810', '22812', '22818', '22819', '22830', '22840', '22841', '22842', '22843', '22844', '22845', '22846', '22847', '22848', '22849', '22850', '22852', '22853', '22854', '22855', '22856', '22857', '22858', '22859', '22861', '22862', '22864', '22865', '22867', '22868', '22869', '22870')
               THEN 1 ELSE 0 END) AS spine,
      MAX(CASE
               WHEN z.PX IN('0WU80JZ', '0WU84JZ', '34.74', '21740', '21742', '21743')
               THEN 1 ELSE 0 END) AS pectus,
      z.PROCDATE,
      YEAR(z.ADATE) as px_date
FROM
         DEMOGRAPHICS a
         INNER JOIN [PROCEDURES] z
              ON z.PERSON_ID = a.PERSON_ID
    WHERE  z.ADATE >= '01/01/2011'
           AND DATEDIFF(YY, a.BIRTH_DATE, z.ADATE) - CASE
                      WHEN DATEADD(YY, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE), a.BIRTH_DATE) > z.ADATE
                      THEN 1
                      ELSE 0
                    END >= 10
           AND DATEDIFF(YY, a.BIRTH_DATE, z.ADATE) - CASE
                      WHEN DATEADD(YY, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE), a.BIRTH_DATE) > z.ADATE
                      THEN 1
                      ELSE 0
                    END <= 26
           AND 
               z.PX IN('0RG4070', '0RG4071', '0RG407J', '0RG40A0', '0RG40A1', '0RG40AJ', '0RG40J0', '0RG40J1', '0RG40JJ', '0RG40K0', '0RG40K1', '0RG40KJ', '0RG40Z0', '0RG40Z1', '0RG40ZJ', '0RG6070', '0RG6071', '0RG607J', '0RG60A0', '0RG60A1', '0RG60AJ', '0RG60J0', '0RG60J1', '0RG60JJ', '0RG60K0', '0RG60K1', '0RG60KJ', '0RG60Z0', '0RG60Z1', '0RG60ZJ', '0RG7070', '0RG7071', '0RG707J', '0RG70A0', '0RG70A1', '0RG70AJ', '0RG70J0', '0RG70J1', '0RG70JJ', '0RG70K0', '0RG70K1', '0RG70KJ', '0RG70Z0', '0RG70Z1', '0RG70ZJ', '0RG8070', '0RG8071', '0RG807J', '0RG80A0', '0RG80A1', '0RG80AJ', '0RG80J0', '0RG80J1', '0RG80JJ', '0RG80K0', '0RG80K1', '0RG80KJ', '0RG80Z0', '0RG80Z1', '0RG80ZJ', '0RGA070', '0RGA071', '0RGA07J', '0RGA0A0', '0RGA0A1', '0RGA0AJ', '0RGA0J0', '0RGA0J1', '0RGA0JJ', '0RGA0K0', '0RGA0K1', '0RGA0KJ', '0RGA0Z0', '0RGA0Z1', '0RGA0ZJ', '0SG0070', '0SG0071', '0SG007J', '0SG00A0', '0SG00A1', '0SG00AJ', '0SG00J0', '0SG00J1', '0SG00JJ', '0SG00K0', '0SG00K1', '0SG00KJ', '0SG00Z0', '0SG00Z1', '0SG00ZJ', '0SG1070', '0SG1071', '0SG107J', '0SG10A0', '0SG10A1', '0SG10AJ', '0SG10J0', '0SG10J1', '0SG10JJ', '0SG10K0', '0SG10K1', '0SG10KJ', '0SG10Z0', '0SG10Z1', '0SG10ZJ', '0SG3070', '0SG3071', '0SG307J', '0SG30A0', '0SG30A1', '0SG30AJ', '0SG30J0', '0SG30J1', '0SG30JJ', '0SG30K0', '0SG30K1', '0SG30KJ', '0SG30Z0', '0SG30Z1', '0SG30ZJ', '81', '81.01', '81.02', '81.03', '81.04', '81.05', '81.06', '81.07', '81.08', '81.09', '81.3', '81.35', '81.36', '81.37', '81.38', '81.39', '81.61', '81.62', '81.63', '81.64', '84.51', '0WU80JZ', '0WU84JZ', '34.74', '22532', '22533', '22548', '22551', '22554', '22556', '22558', '22586', '22590', '22595', '22600', '22610', '22612', '22614', '22630', '22633', '22634', '22800', '22802', '22804', '22808', '22810', '22812', '22818', '22819', '22830', '22840', '22841', '22842', '22843', '22844', '22845', '22846', '22847', '22848', '22849', '22850', '22852', '22853', '22854', '22855', '22856', '22857', '22858', '22859', '22861', '22862', '22864', '22865', '22867', '22868', '22869', '22870', '21740', '21742', '21743')
GROUP BY z.PERSON_ID, z.ENC_ID, z.PX, z.PROCDATE, YEAR(z.ADATE)"

)

tab2b2 <- tab2b %>% 
  group_by(PERSON_ID) %>% 
  mutate(
    min_dt = min(as.Date(PROCDATE)),
    use_dt = if_else(as.Date(min_dt) == as.Date(PROCDATE), 1, 0)
  ) %>% 
  ungroup(.) %>% 
  dplyr::filter(use_dt == 1) %>% 
  left_join(., los_tab, by = "PERSON_ID")

tab2b_test <- as.data.frame(table(tab2b2$ProcedureType, tab2b2$PX))

# list_tab2b <- split(tab2b_test[, c(1,2)], seq(nrow(tab2b_test[, c(1,2)])))
# 
# px_tab <- data.frame(avg_los = unlist(lapply(list_tab2b, 
#                                       function(x) mean(as.numeric(tab2b2[tab2b2$ProcedureType == x[[1]] & tab2b2$PX == x[[2]],
#                                                               "los2"]), na.rm = T))))
# 
# mean(as.numeric(tab2b2[tab2b2$ProcedureType == list_tab2b[[1]][[1]] & tab2b2$PX == list_tab2b[[1]][[2]],
#                                                               "los2"]), na.rm = T)
# 
# px_tab2 <- data.frame(med_los = unlist(lapply(list_tab2b, 
#                                       function(x) median(as.numeric(tab2b2[tab2b2$ProcedureType == x[[1]] & tab2b2$PX == x[[2]],
#                                                               "los2"]), na.rm = T))))
# 
# tab2b3 <- bind_cols(tab2b_test, px_tab)
# tab2b4 <- bind_cols(tab2b3, px_tab2) %>% 
#   dplyr::rename(ProcedureType = Var1, PX = Var2)

```

```{r tab2b_out}
knitr::kable(tab2b_test, col.names = c("Procedure Type", "Procedure", "Count"))

```

## Table 2C: Social History Frequencies by Location


```{r, tab2c_data}

tab2c <- odbc::dbGetQuery(con,
       "SELECT
        DISTINCT z.ENC_ID,
        z.PERSON_ID,
        d.TOBACCO_USER,
        d.ALCOHOL_USER,
        d.ILL_DRUG_USER,
        d.CONDOM_YN,
        d.PILL_YN,
        d.DIAPHRAGM_YN,
        d.IUD_YN,
        d.SURGICAL_YN,
        d.SPERMICIDE_YN,
        d.IMPLANT_YN,
        d.INJECTION_YN,
        d.SPONGE_YN,
        d.BC_INSERTS_YN,
        d.CONTACT_DATE
FROM
         DEMOGRAPHICS a
         INNER JOIN [PROCEDURES] z
              ON z.PERSON_ID = a.PERSON_ID
         LEFT JOIN [SOCIAL_HISTORY] d
              ON z.PERSON_ID = d.PERSON_ID AND z.ENC_ID = d.ENC_ID
    WHERE  z.ADATE >= '01/01/2011'
           AND DATEDIFF(YY, a.BIRTH_DATE, z.ADATE) - CASE
                      WHEN DATEADD(YY, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE), a.BIRTH_DATE) > z.ADATE
                      THEN 1
                      ELSE 0
                    END >= 10
           AND DATEDIFF(YY, a.BIRTH_DATE, z.ADATE) - CASE
                      WHEN DATEADD(YY, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE), a.BIRTH_DATE) > z.ADATE
                      THEN 1
                      ELSE 0
                    END <= 26
           AND 
               z.PX IN('0RG4070', '0RG4071', '0RG407J', '0RG40A0', '0RG40A1', '0RG40AJ', '0RG40J0', '0RG40J1', '0RG40JJ', '0RG40K0', '0RG40K1', '0RG40KJ', '0RG40Z0', '0RG40Z1', '0RG40ZJ', '0RG6070', '0RG6071', '0RG607J', '0RG60A0', '0RG60A1', '0RG60AJ', '0RG60J0', '0RG60J1', '0RG60JJ', '0RG60K0', '0RG60K1', '0RG60KJ', '0RG60Z0', '0RG60Z1', '0RG60ZJ', '0RG7070', '0RG7071', '0RG707J', '0RG70A0', '0RG70A1', '0RG70AJ', '0RG70J0', '0RG70J1', '0RG70JJ', '0RG70K0', '0RG70K1', '0RG70KJ', '0RG70Z0', '0RG70Z1', '0RG70ZJ', '0RG8070', '0RG8071', '0RG807J', '0RG80A0', '0RG80A1', '0RG80AJ', '0RG80J0', '0RG80J1', '0RG80JJ', '0RG80K0', '0RG80K1', '0RG80KJ', '0RG80Z0', '0RG80Z1', '0RG80ZJ', '0RGA070', '0RGA071', '0RGA07J', '0RGA0A0', '0RGA0A1', '0RGA0AJ', '0RGA0J0', '0RGA0J1', '0RGA0JJ', '0RGA0K0', '0RGA0K1', '0RGA0KJ', '0RGA0Z0', '0RGA0Z1', '0RGA0ZJ', '0SG0070', '0SG0071', '0SG007J', '0SG00A0', '0SG00A1', '0SG00AJ', '0SG00J0', '0SG00J1', '0SG00JJ', '0SG00K0', '0SG00K1', '0SG00KJ', '0SG00Z0', '0SG00Z1', '0SG00ZJ', '0SG1070', '0SG1071', '0SG107J', '0SG10A0', '0SG10A1', '0SG10AJ', '0SG10J0', '0SG10J1', '0SG10JJ', '0SG10K0', '0SG10K1', '0SG10KJ', '0SG10Z0', '0SG10Z1', '0SG10ZJ', '0SG3070', '0SG3071', '0SG307J', '0SG30A0', '0SG30A1', '0SG30AJ', '0SG30J0', '0SG30J1', '0SG30JJ', '0SG30K0', '0SG30K1', '0SG30KJ', '0SG30Z0', '0SG30Z1', '0SG30ZJ', '81', '81.01', '81.02', '81.03', '81.04', '81.05', '81.06', '81.07', '81.08', '81.09', '81.3', '81.35', '81.36', '81.37', '81.38', '81.39', '81.61', '81.62', '81.63', '81.64', '84.51', '0WU80JZ', '0WU84JZ', '34.74', '22532', '22533', '22548', '22551', '22554', '22556', '22558', '22586', '22590', '22595', '22600', '22610', '22612', '22614', '22630', '22633', '22634', '22800', '22802', '22804', '22808', '22810', '22812', '22818', '22819', '22830', '22840', '22841', '22842', '22843', '22844', '22845', '22846', '22847', '22848', '22849', '22850', '22852', '22853', '22854', '22855', '22856', '22857', '22858', '22859', '22861', '22862', '22864', '22865', '22867', '22868', '22869', '22870', '21740', '21742', '21743')"
)

# Determine if date of social history assessment was during procedure encounter
tab2c$visit <- ave(tab2c$PERSON_ID == tab2c$PERSON_ID, tab2c$PERSON_ID, FUN = cumsum)

# Get count of how many didn't get asked smoking

tab2c2 <- tab2c %>% 
  left_join(., tab1d, by = c("PERSON_ID", "ENC_ID")) %>% 
  group_by(PERSON_ID) %>% 
  mutate(
    sh_dt   = if_else(CONTACT_DATE >= ADATE & CONTACT_DATE <= DDATE, 1, 0),
    max_vis = max(visit, na.rm = T),
    use_vis = if_else(sh_dt == 1, 1, 
                      if_else(max_vis == 1, 1, 0))
  ) %>%
  dplyr::filter(use_vis == 1) %>% 
  mutate(
    ever_smoke = if_else(TOBACCO_USER == "Y", 1, 0),
    max_smoke  = max(ever_smoke),
    use_smoke  = if_else(ever_smoke == max_smoke, 1, 0),
    bc_use     = if_else(CONDOM_YN == "Y" | 
        PILL_YN == "Y" |
        DIAPHRAGM_YN == "Y" |
        IUD_YN == "Y" |
        SURGICAL_YN == "Y" |
        SPERMICIDE_YN == "Y" |
        IMPLANT_YN == "Y" |
        INJECTION_YN == "Y" |
        SPONGE_YN == "Y" |
        BC_INSERTS_YN == "Y", 1, 0),
    max_bc    = max(bc_use, na.rm = T),
    alc_use   = if_else(ALCOHOL_USER == "Y", 1, 0),
    max_alc   = max(alc_use, na.rm = T),
    drug_use  = if_else(ILL_DRUG_USER == "Y", 1, 0)
  ) %>% 
  dplyr::filter(use_smoke == 1 & TOBACCO_USER != "X") %>% 
  dplyr::filter(!duplicated(PERSON_ID)) %>% 
  ungroup(.)

mean_los_sh <- c(mean(tab2c2$los2[tab2c2$max_smoke == 1], na.rm = T),
                 mean(tab2c2$los2[tab2c2$max_alc == 1], na.rm = T),
                 mean(tab2c2$los2[tab2c2$drug_use == 1], na.rm = T),
                 mean(tab2c2$los2[tab2c2$max_bc == 1], na.rm = T))

median_los_sh <- c(median(tab2c2$los2[tab2c2$max_smoke == 1], na.rm = T),
                   median(tab2c2$los2[tab2c2$max_alc == 1], na.rm = T),
                   median(tab2c2$los2[tab2c2$drug_use == 1], na.rm = T),
                   median(tab2c2$los2[tab2c2$max_bc == 1], na.rm = T))
count_sh <- c(tally(tab2c2[tab2c2$max_smoke == 1, "los2"])[[1]],
              tally(tab2c2[tab2c2$max_alc == 1, "los2"])[[1]],
              tally(tab2c2[tab2c2$drug_use == 1, "los2"])[[1]],
              tally(tab2c2[tab2c2$max_bc == 1, "los2"])[[1]])

names_sh <- c("Tobacco Use", "Alcohol Use", "Illegal Drug Use",  "Birth Control Use")

tab2c_sum <- data.frame(names = names_sh, counts = count_sh, means = mean_los_sh, medians = median_los_sh)

```

```{r tab2c_out, results="asis"}

knitr::kable(tab2c_sum, col.names = c(paste0("Variables N = ", nrow(tab2c2)), "Count", "Mean LOS", "Median LOS")) 

```

# Data Tables Protocol: Table Set 3: Provider Information

## Table 3A: Type of Surgeon/Physician of Record for Index

This table contains information only for physicians, not fellows, residents, or other providers.

```{r med_names}

tab3a <- odbc::dbGetQuery(con,
       "SELECT
        DISTINCT z.ENC_ID,
        z.PERSON_ID,
        z.PX,
        z.PROVIDER,
        z.PERFORMINGPROVIDER,
        d.SPECIALTY,
        d.SPECIALTY2,
        d.SPECIALTY3,
        d.SPECIALTY4,
        d.PROVIDER_TYPE,
       MAX(CASE.
               WHEN z.PX IN('0RG4070', '0RG4071', '0RG407J', '0RG40A0', '0RG40A1', '0RG40AJ', '0RG40J0', '0RG40J1', '0RG40JJ', '0RG40K0', '0RG40K1', '0RG40KJ', '0RG40Z0', '0RG40Z1', '0RG40ZJ', '0RG6070', '0RG6071', '0RG607J', '0RG60A0', '0RG60A1', '0RG60AJ', '0RG60J0', '0RG60J1', '0RG60JJ', '0RG60K0', '0RG60K1', '0RG60KJ', '0RG60Z0', '0RG60Z1', '0RG60ZJ', '0RG7070', '0RG7071', '0RG707J', '0RG70A0', '0RG70A1', '0RG70AJ', '0RG70J0', '0RG70J1', '0RG70JJ', '0RG70K0', '0RG70K1', '0RG70KJ', '0RG70Z0', '0RG70Z1', '0RG70ZJ', '0RG8070', '0RG8071', '0RG807J', '0RG80A0', '0RG80A1', '0RG80AJ', '0RG80J0', '0RG80J1', '0RG80JJ', '0RG80K0', '0RG80K1', '0RG80KJ', '0RG80Z0', '0RG80Z1', '0RG80ZJ', '0RGA070', '0RGA071', '0RGA07J', '0RGA0A0', '0RGA0A1', '0RGA0AJ', '0RGA0J0', '0RGA0J1', '0RGA0JJ', '0RGA0K0', '0RGA0K1', '0RGA0KJ', '0RGA0Z0', '0RGA0Z1', '0RGA0ZJ', '0SG0070', '0SG0071', '0SG007J', '0SG00A0', '0SG00A1', '0SG00AJ', '0SG00J0', '0SG00J1', '0SG00JJ', '0SG00K0', '0SG00K1', '0SG00KJ', '0SG00Z0', '0SG00Z1', '0SG00ZJ', '0SG1070', '0SG1071', '0SG107J', '0SG10A0', '0SG10A1', '0SG10AJ', '0SG10J0', '0SG10J1', '0SG10JJ', '0SG10K0', '0SG10K1', '0SG10KJ', '0SG10Z0', '0SG10Z1', '0SG10ZJ', '0SG3070', '0SG3071', '0SG307J', '0SG30A0', '0SG30A1', '0SG30AJ', '0SG30J0', '0SG30J1', '0SG30JJ', '0SG30K0', '0SG30K1', '0SG30KJ', '0SG30Z0', '0SG30Z1', '0SG30ZJ', '81', '81.01', '81.02', '81.03', '81.04', '81.05', '81.06', '81.07', '81.08', '81.09', '81.3', '81.35', '81.36', '81.37', '81.38', '81.39', '81.61', '81.62', '81.63', '81.64', '84.51', '22532', '22533', '22548', '22551', '22554', '22556', '22558', '22586', '22590', '22595', '22600', '22610', '22612', '22614', '22630', '22633', '22634', '22800', '22802', '22804', '22808', '22810', '22812', '22818', '22819', '22830', '22840', '22841', '22842', '22843', '22844', '22845', '22846', '22847', '22848', '22849', '22850', '22852', '22853', '22854', '22855', '22856', '22857', '22858', '22859', '22861', '22862', '22864', '22865', '22867', '22868', '22869', '22870')
               THEN 1 ELSE 0 END) AS spine,
      MAX(CASE
               WHEN z.PX IN('0WU80JZ', '0WU84JZ', '34.74', '21740', '21742', '21743')
               THEN 1 ELSE 0 END) AS pectus,
      z.PROCDATE,
      YEAR(z.ADATE) as px_date
FROM
         DEMOGRAPHICS a
         INNER JOIN [PROCEDURES] z
              ON z.PERSON_ID = a.PERSON_ID
         LEFT JOIN [PROVIDER_SPECIALTY] d
              ON z.PERFORMINGPROVIDER = d.PROVIDER
    WHERE  z.ADATE >= '01/01/2011'
           AND DATEDIFF(YY, a.BIRTH_DATE, z.ADATE) - CASE
                      WHEN DATEADD(YY, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE), a.BIRTH_DATE) > z.ADATE
                      THEN 1
                      ELSE 0
                    END >= 10
           AND DATEDIFF(YY, a.BIRTH_DATE, z.ADATE) - CASE
                      WHEN DATEADD(YY, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE), a.BIRTH_DATE) > z.ADATE
                      THEN 1
                      ELSE 0
                    END <= 26
           AND 
               z.PX IN('0RG4070', '0RG4071', '0RG407J', '0RG40A0', '0RG40A1', '0RG40AJ', '0RG40J0', '0RG40J1', '0RG40JJ', '0RG40K0', '0RG40K1', '0RG40KJ', '0RG40Z0', '0RG40Z1', '0RG40ZJ', '0RG6070', '0RG6071', '0RG607J', '0RG60A0', '0RG60A1', '0RG60AJ', '0RG60J0', '0RG60J1', '0RG60JJ', '0RG60K0', '0RG60K1', '0RG60KJ', '0RG60Z0', '0RG60Z1', '0RG60ZJ', '0RG7070', '0RG7071', '0RG707J', '0RG70A0', '0RG70A1', '0RG70AJ', '0RG70J0', '0RG70J1', '0RG70JJ', '0RG70K0', '0RG70K1', '0RG70KJ', '0RG70Z0', '0RG70Z1', '0RG70ZJ', '0RG8070', '0RG8071', '0RG807J', '0RG80A0', '0RG80A1', '0RG80AJ', '0RG80J0', '0RG80J1', '0RG80JJ', '0RG80K0', '0RG80K1', '0RG80KJ', '0RG80Z0', '0RG80Z1', '0RG80ZJ', '0RGA070', '0RGA071', '0RGA07J', '0RGA0A0', '0RGA0A1', '0RGA0AJ', '0RGA0J0', '0RGA0J1', '0RGA0JJ', '0RGA0K0', '0RGA0K1', '0RGA0KJ', '0RGA0Z0', '0RGA0Z1', '0RGA0ZJ', '0SG0070', '0SG0071', '0SG007J', '0SG00A0', '0SG00A1', '0SG00AJ', '0SG00J0', '0SG00J1', '0SG00JJ', '0SG00K0', '0SG00K1', '0SG00KJ', '0SG00Z0', '0SG00Z1', '0SG00ZJ', '0SG1070', '0SG1071', '0SG107J', '0SG10A0', '0SG10A1', '0SG10AJ', '0SG10J0', '0SG10J1', '0SG10JJ', '0SG10K0', '0SG10K1', '0SG10KJ', '0SG10Z0', '0SG10Z1', '0SG10ZJ', '0SG3070', '0SG3071', '0SG307J', '0SG30A0', '0SG30A1', '0SG30AJ', '0SG30J0', '0SG30J1', '0SG30JJ', '0SG30K0', '0SG30K1', '0SG30KJ', '0SG30Z0', '0SG30Z1', '0SG30ZJ', '81', '81.01', '81.02', '81.03', '81.04', '81.05', '81.06', '81.07', '81.08', '81.09', '81.3', '81.35', '81.36', '81.37', '81.38', '81.39', '81.61', '81.62', '81.63', '81.64', '84.51', '0WU80JZ', '0WU84JZ', '34.74', '22532', '22533', '22548', '22551', '22554', '22556', '22558', '22586', '22590', '22595', '22600', '22610', '22612', '22614', '22630', '22633', '22634', '22800', '22802', '22804', '22808', '22810', '22812', '22818', '22819', '22830', '22840', '22841', '22842', '22843', '22844', '22845', '22846', '22847', '22848', '22849', '22850', '22852', '22853', '22854', '22855', '22856', '22857', '22858', '22859', '22861', '22862', '22864', '22865', '22867', '22868', '22869', '22870', '21740', '21742', '21743')
GROUP BY z.PERSON_ID, z.ENC_ID, z.PX, z.PROVIDER,
        z.PERFORMINGPROVIDER,
        d.SPECIALTY,
        d.SPECIALTY2,
        d.SPECIALTY3,
        d.SPECIALTY4,
        d.PROVIDER_TYPE, z.PROCDATE, YEAR(z.ADATE)"
)

tab3a_prov <- odbc::dbGetQuery(con,
       "SELECT
        z.ENC_ID,
        z.PERSON_ID,
        z.PX,
        z.PROVIDER,
        z.PERFORMINGPROVIDER,
        d.SPECIALTY,
        d.SPECIALTY2,
        d.SPECIALTY3,
        d.SPECIALTY4,
        d.PROVIDER_TYPE,
       MAX(CASE
               WHEN z.PX IN('0RG4070', '0RG4071', '0RG407J', '0RG40A0', '0RG40A1', '0RG40AJ', '0RG40J0', '0RG40J1', '0RG40JJ', '0RG40K0', '0RG40K1', '0RG40KJ', '0RG40Z0', '0RG40Z1', '0RG40ZJ', '0RG6070', '0RG6071', '0RG607J', '0RG60A0', '0RG60A1', '0RG60AJ', '0RG60J0', '0RG60J1', '0RG60JJ', '0RG60K0', '0RG60K1', '0RG60KJ', '0RG60Z0', '0RG60Z1', '0RG60ZJ', '0RG7070', '0RG7071', '0RG707J', '0RG70A0', '0RG70A1', '0RG70AJ', '0RG70J0', '0RG70J1', '0RG70JJ', '0RG70K0', '0RG70K1', '0RG70KJ', '0RG70Z0', '0RG70Z1', '0RG70ZJ', '0RG8070', '0RG8071', '0RG807J', '0RG80A0', '0RG80A1', '0RG80AJ', '0RG80J0', '0RG80J1', '0RG80JJ', '0RG80K0', '0RG80K1', '0RG80KJ', '0RG80Z0', '0RG80Z1', '0RG80ZJ', '0RGA070', '0RGA071', '0RGA07J', '0RGA0A0', '0RGA0A1', '0RGA0AJ', '0RGA0J0', '0RGA0J1', '0RGA0JJ', '0RGA0K0', '0RGA0K1', '0RGA0KJ', '0RGA0Z0', '0RGA0Z1', '0RGA0ZJ', '0SG0070', '0SG0071', '0SG007J', '0SG00A0', '0SG00A1', '0SG00AJ', '0SG00J0', '0SG00J1', '0SG00JJ', '0SG00K0', '0SG00K1', '0SG00KJ', '0SG00Z0', '0SG00Z1', '0SG00ZJ', '0SG1070', '0SG1071', '0SG107J', '0SG10A0', '0SG10A1', '0SG10AJ', '0SG10J0', '0SG10J1', '0SG10JJ', '0SG10K0', '0SG10K1', '0SG10KJ', '0SG10Z0', '0SG10Z1', '0SG10ZJ', '0SG3070', '0SG3071', '0SG307J', '0SG30A0', '0SG30A1', '0SG30AJ', '0SG30J0', '0SG30J1', '0SG30JJ', '0SG30K0', '0SG30K1', '0SG30KJ', '0SG30Z0', '0SG30Z1', '0SG30ZJ', '81', '81.01', '81.02', '81.03', '81.04', '81.05', '81.06', '81.07', '81.08', '81.09', '81.3', '81.35', '81.36', '81.37', '81.38', '81.39', '81.61', '81.62', '81.63', '81.64', '84.51', '22532', '22533', '22548', '22551', '22554', '22556', '22558', '22586', '22590', '22595', '22600', '22610', '22612', '22614', '22630', '22633', '22634', '22800', '22802', '22804', '22808', '22810', '22812', '22818', '22819', '22830', '22840', '22841', '22842', '22843', '22844', '22845', '22846', '22847', '22848', '22849', '22850', '22852', '22853', '22854', '22855', '22856', '22857', '22858', '22859', '22861', '22862', '22864', '22865', '22867', '22868', '22869', '22870')
               THEN 1 ELSE 0 END) AS spine,
      MAX(CASE
               WHEN z.PX IN('0WU80JZ', '0WU84JZ', '34.74', '21740', '21742', '21743')
               THEN 1 ELSE 0 END) AS pectus,
      z.PROCDATE,
      YEAR(z.ADATE) as px_date
FROM
         DEMOGRAPHICS a
         INNER JOIN [PROCEDURES] z
              ON z.PERSON_ID = a.PERSON_ID
         LEFT JOIN [PROVIDER_SPECIALTY] d
              ON z.PERFORMINGPROVIDER = d.PROVIDER
    WHERE  z.ADATE >= '01/01/2011'
           AND DATEDIFF(YY, a.BIRTH_DATE, z.ADATE) - CASE
                      WHEN DATEADD(YY, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE), a.BIRTH_DATE) > z.ADATE
                      THEN 1
                      ELSE 0
                    END >= 10
           AND DATEDIFF(YY, a.BIRTH_DATE, z.ADATE) - CASE
                      WHEN DATEADD(YY, DATEDIFF(YY, a.BIRTH_DATE, z.ADATE), a.BIRTH_DATE) > z.ADATE
                      THEN 1
                      ELSE 0
                    END <= 26
           AND 
               z.PERFORMINGPROVIDER IN('E_13066', 'E_13066', 'E_13066', 'E_13066', 'E_13066', 'E_13066', 'E_13066', '088377', 'E_13066', '088377', 'E_13066', 'E_13066', 'E_13066', 'E_1526', 'E_1526', 'E_1322', 'E_1322', 'E_1322', 'E_1322', 'E_1322', 'E_13066', 'E_13066', 'E_13066', 'E_13066', 'E_1322', 'E_1322', 'E_1322', 'E_1322', 'E_1322', 'E_13066', 'E_18112', 'E_13066', 'E_18112', 'E_13066', 'E_18112', '116038', '116038', '158352', '158352', '158352', '145318', '145466', '145466', '145466', '145466', '145466', '145466', '145466', '123364', '123364', '130666', '130666', '169649', '169649', '169649', '125401', '125401', '125401', '086330', '086330', '086330', '086330', '086330', 'E_1322', 'E_1322', '132274', '132274', 'E_13066', 'E_13066', '113399', '113399', '073361', '073361', '073361', '116038', '116038', '116038', '116038', '116038', '130666', '123364', '123364', '123364', '123364', '133595', '133595', '133595', '145318', '140848', '140848', '140848', '125401', '125401', '125401', 'E_13066', 'E_13066', 'E_13066', 'E_13066', '125401', '125401', '125401', '125401', '124586', '124586', '124586', 'E_13066', 'E_13066', 'E_13066', 'E_13066', '073361', '073361', '073361', '128850', '128850', '124453', '124453', '158352', '158352', '158352', '158352', '158352', '100305', '100305', '141531', '141531', '141531', '132274', '132274', '132274', '046771', '046771', '046771', 'E_13066', '133322', '133322', '133322', '125401', '125401', '125401', '141531', '141531', '141531', '070979', '070979', '130666', '130666', '130666', '130666', '130666', '130666', '123364', '124453', '124453', '158378', '158378', '158378', '158378', '131375', '131375', '131375', '133595', '133595', '133322', '133322', '086330', '086330', '086330', '086330', '086330', 'E_1526', 'E_1526', 'E_1526', '133322', '133322', '156604', '156604', 'E_13066', '088377', '088377', '088377', '141531', '141531', '079038', '079038', '079038', '079038', '079038', '077321', '157230', '157230', '100305', '100305', '100305', '100305', '100305', '141440', '141440', '133322', '133322', '073411', '073411', '073411', '133595', '133595', '128850', '128850', '156604', '156604', '156604', '156604', '156604', '156604', '128850', '128850', 'E_1526', 'E_1526', 'E_1526', 'E_1526', 'E_1526', '123364', '123364', '123364', '158378', '158378', '124453', '124453', '124453', '156604', '156604', '141465', '141465', '141465', '124453', '124453', '124453', '124453', '128850', '128850', '128850', '128850', '128850', '146787', '146787', '119578', '119578', '147918', '147918', '147918', '124586', '086330', '086330', '086330', '153767', '153767', '110676', '110676', '111153', '111153') 
            AND
               z.PX IN('0RG4070', '0RG4071', '0RG407J', '0RG40A0', '0RG40A1', '0RG40AJ', '0RG40J0', '0RG40J1', '0RG40JJ', '0RG40K0', '0RG40K1', '0RG40KJ', '0RG40Z0', '0RG40Z1', '0RG40ZJ', '0RG6070', '0RG6071', '0RG607J', '0RG60A0', '0RG60A1', '0RG60AJ', '0RG60J0', '0RG60J1', '0RG60JJ', '0RG60K0', '0RG60K1', '0RG60KJ', '0RG60Z0', '0RG60Z1', '0RG60ZJ', '0RG7070', '0RG7071', '0RG707J', '0RG70A0', '0RG70A1', '0RG70AJ', '0RG70J0', '0RG70J1', '0RG70JJ', '0RG70K0', '0RG70K1', '0RG70KJ', '0RG70Z0', '0RG70Z1', '0RG70ZJ', '0RG8070', '0RG8071', '0RG807J', '0RG80A0', '0RG80A1', '0RG80AJ', '0RG80J0', '0RG80J1', '0RG80JJ', '0RG80K0', '0RG80K1', '0RG80KJ', '0RG80Z0', '0RG80Z1', '0RG80ZJ', '0RGA070', '0RGA071', '0RGA07J', '0RGA0A0', '0RGA0A1', '0RGA0AJ', '0RGA0J0', '0RGA0J1', '0RGA0JJ', '0RGA0K0', '0RGA0K1', '0RGA0KJ', '0RGA0Z0', '0RGA0Z1', '0RGA0ZJ', '0SG0070', '0SG0071', '0SG007J', '0SG00A0', '0SG00A1', '0SG00AJ', '0SG00J0', '0SG00J1', '0SG00JJ', '0SG00K0', '0SG00K1', '0SG00KJ', '0SG00Z0', '0SG00Z1', '0SG00ZJ', '0SG1070', '0SG1071', '0SG107J', '0SG10A0', '0SG10A1', '0SG10AJ', '0SG10J0', '0SG10J1', '0SG10JJ', '0SG10K0', '0SG10K1', '0SG10KJ', '0SG10Z0', '0SG10Z1', '0SG10ZJ', '0SG3070', '0SG3071', '0SG307J', '0SG30A0', '0SG30A1', '0SG30AJ', '0SG30J0', '0SG30J1', '0SG30JJ', '0SG30K0', '0SG30K1', '0SG30KJ', '0SG30Z0', '0SG30Z1', '0SG30ZJ', '81', '81.01', '81.02', '81.03', '81.04', '81.05', '81.06', '81.07', '81.08', '81.09', '81.3', '81.35', '81.36', '81.37', '81.38', '81.39', '81.61', '81.62', '81.63', '81.64', '84.51', '0WU80JZ', '0WU84JZ', '34.74', '22532', '22533', '22548', '22551', '22554', '22556', '22558', '22586', '22590', '22595', '22600', '22610', '22612', '22614', '22630', '22633', '22634', '22800', '22802', '22804', '22808', '22810', '22812', '22818', '22819', '22830', '22840', '22841', '22842', '22843', '22844', '22845', '22846', '22847', '22848', '22849', '22850', '22852', '22853', '22854', '22855', '22856', '22857', '22858', '22859', '22861', '22862', '22864', '22865', '22867', '22868', '22869', '22870', '21740', '21742', '21743')
GROUP BY z.ENC_ID,
        z.PERSON_ID,
        z.PX,
        z.PROVIDER,
        z.PERFORMINGPROVIDER,
        d.SPECIALTY,
        d.SPECIALTY2,
        d.SPECIALTY3,
        d.SPECIALTY4,
        d.PROVIDER_TYPE, z.PROCDATE, YEAR(z.ADATE)"
)

tab3a2_prov <- tab3a_prov %>% 
  dplyr::filter(!duplicated(ENC_ID)) %>% 
  group_by(PERFORMINGPROVIDER, px_date) %>% 
  summarise(n_procs = n(), spine = sum(spine), pectus = sum(pectus))

tab3a2 <- tab3a %>%
  left_join(., specialties, by = "SPECIALTY") %>% 
  dplyr::filter(PROVIDER_TYPE %in% c(61, 25, 77, 999)) %>% 
  dplyr::filter(!duplicated(ENC_ID))

```


```{r tab2_res}
tab2_lab_null <- odbc::dbGetQuery(con, 
  "SELECT SUM(CASE WHEN TEST_TYPE IS NULL THEN 1 ELSE 0 END) as TEST_TYPE,
   SUM(CASE WHEN BATTERY_CD IS NULL THEN 1 ELSE 0 END) as BATTERY_CD,
   SUM(CASE WHEN LOINC IS NULL THEN 1 ELSE 0 END) as LOINC,
   SUM(CASE WHEN LOCAL_CD IS NULL THEN 1 ELSE 0 END) as LOCAL_CD,
   SUM(CASE WHEN RESULT_NUM IS NULL THEN 1 ELSE 0 END) as RESULT_NUM,
   SUM(CASE WHEN RESULT_UNIT IS NULL THEN 1 ELSE 0 END) as RESULT_UNIT,
   SUM(CASE WHEN SPECIMEN_SOURCE IS NULL THEN 1 ELSE 0 END) as SPECIMEN_SOURCE
      from lab_results")

tab2_lab_null2 <- data.frame(table = rep("LAB_RESULTS", ncol(tab2_lab_null)), var = names(tab2_lab_null), records_null = t(tab2_lab_null)[,1])
row.names(tab2_lab_null2) <- NULL

tab2_rx_null <- odbc::dbGetQuery(con, 
  "SELECT SUM(CASE WHEN GENERIC_MED_NAME IS NULL THEN 1 ELSE 0 END) as GENERIC_MED_NAME,
   SUM(CASE WHEN RX_DAYS_SUPPLY IS NULL THEN 1 ELSE 0 END) as RX_DAYS_SUPPLY,
   SUM(CASE WHEN RX_END_DATE IS NULL THEN 1 ELSE 0 END) as RX_END_DATE,
   SUM(CASE WHEN RX_START_DATE IS NULL THEN 1 ELSE 0 END) as RX_START_DATE,
   SUM(CASE WHEN RX_QUANTITY IS NULL THEN 1 ELSE 0 END) as RX_QUANTITY,
   SUM(CASE WHEN RX_QUANTITY_NUM IS NULL THEN 1 ELSE 0 END) as RX_QUANTITY_NUM,
   SUM(CASE WHEN RX_QUANTITY_UNIT IS NULL THEN 1 ELSE 0 END) as RX_QUANTITY_UNIT,
   SUM(CASE WHEN RX_REFILLS IS NULL THEN 1 ELSE 0 END) as RX_REFILLS,
   SUM(CASE WHEN RXMD IS NULL THEN 1 ELSE 0 END) as RXMD,
   SUM(CASE WHEN RXNORM IS NULL THEN 1 ELSE 0 END) as RXNORM
      from prescribing")

tab2_rx_null2 <- data.frame(table = rep("PRESCRIBING", ncol(tab2_rx_null)), var = names(tab2_rx_null), records_null = t(tab2_rx_null)[,1])
row.names(tab2_rx_null2) <- NULL

tab2_px_null <- odbc::dbGetQuery(con, 
  "SELECT SUM(CASE WHEN PROVIDER IS NULL THEN 1 ELSE 0 END) as PROVIDER,
   SUM(CASE WHEN ENCTYPE IS NULL THEN 1 ELSE 0 END) as ENCTYPE,
   SUM(CASE WHEN PXCNT IS NULL THEN 1 ELSE 0 END) as PXCNT
      from procedures")

tab2_px_null2 <- data.frame(table = rep("PROCEDURES", ncol(tab2_px_null)), var = names(tab2_px_null), records_null = t(tab2_px_null)[,1])
row.names(tab2_px_null2) <- NULL

tab2_ps_null <- odbc::dbGetQuery(con, 
  "SELECT SUM(CASE WHEN SPECIALTY IS NULL THEN 1 ELSE 0 END) as SPECIALTY,
   SUM(CASE WHEN PROVIDER_TYPE IS NULL THEN 1 ELSE 0 END) as PROVIDER_TYPE,
   SUM(CASE WHEN PROVIDER_BIRTH_YEAR IS NULL THEN 1 ELSE 0 END) as PROVIDER_BIRTH_YEAR,
   SUM(CASE WHEN PROVIDER_GENDER IS NULL THEN 1 ELSE 0 END) as PROVIDER_GENDER,
   SUM(CASE WHEN PROVIDER_RACE IS NULL THEN 1 ELSE 0 END) as PROVIDER_RACE,
   SUM(CASE WHEN PROVIDER_HISPANIC IS NULL THEN 1 ELSE 0 END) as PROVIDER_HISPANIC
      from provider_specialty")

tab2_ps_null2 <- data.frame(table = rep("PROVIDER_SPECIALTY", ncol(tab2_ps_null)), var = names(tab2_ps_null), records_null = t(tab2_ps_null)[,1])
row.names(tab2_ps_null2) <- NULL

tab2_sh_null <- odbc::dbGetQuery(con, 
  "SELECT SUM(CASE WHEN TOBACCO_USER IS NULL THEN 1 ELSE 0 END) as TOBACCO_USER,
   SUM(CASE WHEN ALCOHOL_USER IS NULL THEN 1 ELSE 0 END) as ALCOHOL_USER,
   SUM(CASE WHEN ILL_DRUG_USER IS NULL THEN 1 ELSE 0 END) as ILL_DRUG_USER,
   SUM(CASE WHEN ENC_ID IS NULL THEN 1 ELSE 0 END) as ENC_ID
      from social_history")

tab2_sh_null2 <- data.frame(table = rep("SOCIAL_HISTORY", ncol(tab2_sh_null)), var = names(tab2_sh_null), records_null = t(tab2_sh_null)[,1])
row.names(tab2_sh_null2) <- NULL

tab2_null <- bind_rows(tab2_lab_null2, tab2_rx_null2, tab2_px_null2, tab2_ps_null2, tab2_sh_null2) %>% 
  mutate(
    pct_rec = ifelse(table == "LAB_RESULTS", (records_null / tab1_lab$nrows * 100), 
                     ifelse(table == "PRESCRIBING", (records_null / tab1_rx$nrows * 100), 
                             ifelse(table == "PROCEDURES", (records_null / tab1_px$nrows * 100), 
                                     ifelse(table == "PROVIDER_SPECIALTY", (records_null / tab1_ps$nrows * 100), 
                                             ifelse(table == "SOCIAL_HISTORY", (records_null / tab1_sh$nrows * 100), NA)))))
  )

tab2_lab_uk <- odbc::dbGetQuery(con, 
  "SELECT SUM(CASE WHEN TEST_TYPE = 'Unknown' THEN 1 ELSE 0 END) as TEST_TYPE
      from lab_results")

tab2_lab_uk2 <- data.frame(table = rep("LAB_RESULTS", ncol(tab2_lab_uk)), var = names(tab2_lab_uk), records_uk = t(tab2_lab_uk)[,1])
row.names(tab2_lab_uk2) <- NULL


tab2_ps_uk <- odbc::dbGetQuery(con, 
  "SELECT SUM(CASE WHEN SPECIALTY = 'UNK' THEN 1 ELSE 0 END) as SPECIALTY,
   SUM(CASE WHEN PROVIDER_TYPE = 999 THEN 1 ELSE 0 END) as PROVIDER_TYPE,
   SUM(CASE WHEN PROVIDER_GENDER = 'U' THEN 1 ELSE 0 END) as PROVIDER_GENDER,
   SUM(CASE WHEN PROVIDER_RACE = 'U' THEN 1 ELSE 0 END) as PROVIDER_RACE,
   SUM(CASE WHEN PROVIDER_HISPANIC = 'U' THEN 1 ELSE 0 END) as PROVIDER_HISPANIC
      from provider_specialty")

tab2_ps_uk2 <- data.frame(table = rep("PROVIDER_SPECIALTY", ncol(tab2_ps_uk)), var = names(tab2_ps_uk), records_uk = t(tab2_ps_uk)[,1])
row.names(tab2_ps_uk2) <- NULL

tab2_sh_uk <- odbc::dbGetQuery(con, 
  "SELECT SUM(CASE WHEN TOBACCO_USER = 'U' THEN 1 ELSE 0 END) as TOBACCO_USER,
   SUM(CASE WHEN ALCOHOL_USER = 'U' THEN 1 ELSE 0 END) as ALCOHOL_USER,
   SUM(CASE WHEN ILL_DRUG_USER = 'U' THEN 1 ELSE 0 END) as ILL_DRUG_USER
      from social_history")

tab2_sh_uk2 <- data.frame(table = rep("SOCIAL_HISTORY", ncol(tab2_sh_uk)), var = names(tab2_sh_uk), records_uk = t(tab2_sh_uk)[,1])
row.names(tab2_sh_uk2) <- NULL

tab2_uk <- bind_rows(tab2_lab_uk2, tab2_ps_uk2, tab2_sh_uk2) %>% 
  mutate(
    pct_rec_uk = ifelse(table == "LAB_RESULTS", (records_uk / tab1_lab$nrows * 100), 
                                     ifelse(table == "PROVIDER_SPECIALTY", (records_uk / tab1_ps$nrows * 100), 
                                             ifelse(table == "SOCIAL_HISTORY", (records_uk / tab1_sh$nrows * 100), NA)))
  )

tab2_all <- left_join(tab2_null, tab2_uk, by = c("table", "var"))

```

```{r tab2_disp, results="asis"}

kable(tab2_all, col.names = c("Table", "Variable", "Null Records", "Percent of Records Null", "Unknown Records", "Percent of Records Unknown"), digits = 2, format.args = list(big.mark = ","))
```

##Examination of the LAB_RESULTS Table 

The following tables explore the contents of the LAB_RESULTS table with a focus on trends over time,CHORDS priority lab results, and most common types of lab tests. 


##Table 3: Annual Trends in the Total Number of Recorded Lab Tests, the Number of Patients Tested and the Percent Change from the Previous Year, by Calendar Year

This table displays trends in the number of recorded lab tests and the number of unique patients tested by year. The percent change in both count values is included to identify unexpected temporal fluctuations. 

```{r tab3_res}

tab3 <- odbc::dbGetQuery(con, 
  "select year(result_dt) as year,
    count(*) as nrows, 
	  count(distinct person_id) as npats
	  from lab_results
	    group by year(result_dt)
      order by year(result_dt)")

tab3 <- tab3 %>% 
  mutate(
    ptc_rec = if_else(is.na(nrows / dplyr::lag(nrows)), "--", paste0(round(nrows / dplyr::lag(nrows) * 100, 0), "%")),
    ptc_pat = if_else(is.na(npats / dplyr::lag(npats)), "--", paste0(round(npats / dplyr::lag(npats) * 100, 0), "%"))
  )

tab3$year <- as.character(tab3$year)

tab3_ord <- tab3[, c(1,2,4,3,5)]

```

```{r tab3_disp, results="asis"}

kable(tab3_ord, col.names = c("Year", "Number of Recorded Lab Tests", "Percent Change From Previous Year", "Number of Patients Tested", "Percent Change From Previous Year"), format.args = list(big.mark = ","))

```


##Table 4: CHORDS Priority Lab Tests - Number/Percentage of Tests, Number/Percentage of Patients Tested, Mean Result Values and Result Ranges, by Select TEST_TYPE Values 

This table illustrates the number (and percent) of priority lab test records including the percent of all lab test records , as well as the number (and percent)  of unique patients tested for select TEST_TYPE values. 

```{r tab4_res}

tab4 <- odbc::dbGetQuery(con, 
  "select test_type,
    count(*) as nrows, 
	  count(distinct person_id) as npats,
    avg(result_num) as avg_resval,
    min(result_num) as min_resval,
    max(result_num) as max_resval
	  from lab_results
	    group by test_type
      order by test_type")

tab4_tot <- data.frame(test_type = "Total", nrows = tab1_lab$nrows, npats = tab1_lab$npats, avg_resval = NA, min_resval = NA, max_resval = NA)
tab4 <- tab4 %>% 
  mutate(
    pct_rec = round(nrows / tab1_lab$nrows * 100, 1) ,
    pct_pat = round(npats / tab1_lab$npats * 100, 1)
  ) %>% 
  bind_rows(., tab4_tot)

tab4$test_type[1] <- "UNKNOWN TEST TYPE"
  
tab4_ord <- tab4[, c(1,2,7,3,8,4,5,6)]  
```

```{r tab4_disp, results="asis"}
kable(tab4_ord, col.names = c("Test Type", "Number of Recorded Lab Tests", "Percentage of All Recorded Tests", "Number of Unique Patients Tested", "Percentage of All Patients", "Mean Result Value", "Minimum Result Value", "Maximum Result Value"), digits = 2, format.args = list(big.mark = ",")) 
```

##Figures 1a-1i: Temporal Trends in the Number of CHORDS Priority Lab Tests, by Select TEST_TYPE Values

These figures shows temporal trends in the number of records for CHORDS priority lab tests.


```{r fig1_res, fig.height=10, fig.width=15}

fig1 <- odbc::dbGetQuery(con, 
  "select test_type, 
    month(result_dt) as month,
    year(result_dt) as year,
    count(*) as nrows
	  from lab_results
      where test_type in ('HGBA1C', 'PREGNANCY', 'HDL', 'LDL', 'TRIGLYCERIDE', 'GLUCOSE', 'CREATININE', 'HEP_C_RNA', 'HEP_C_ANTIBODY', 'HEP_C_GENOTYPE', 'IGE', 'EOSINOPHIL')
	    group by year(result_dt), month(result_dt), test_type") 

fig1_sub <- fig1 %>% 
  arrange(year, month, test_type) %>% 
  mutate(
    date = as.Date(paste(year, month, "01", sep = "-"))
  ) %>% 
  select(-month, -year)

fig1_wide <- spread(fig1_sub, key = test_type, value = nrows)

lapply(2:ncol(fig1_wide), function(x) 
  ggplot(data = fig1_wide, aes(x = date, y = fig1_wide[[x]])) +
         geom_line() +
    ggtitle(names(fig1_wide)[x]) +
    theme(axis.text.x = element_text(angle=45, margin(b = .5), vjust = .5)) +
    scale_x_date(date_labels = "%b %y",date_breaks='1 year') +
    xlab("Date") + ylab("Number of Records") +
    theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),  
        axis.line = element_line(colour = "black"),
        axis.title = element_text(size = 16, face = "bold"),
        axis.text = element_text(size = 14, face = "bold"),
        plot.title = element_text(hjust = 0.5, size = 20))
    ) 


```

##Table 5: Summary of the Ten Most Common Lab Tests - Number/Percentage of Tests and Number/Percentage of Patients Tested, by Select BATTERY_CD Values

This table illustrates the number (and percent) of the data partner's ten most common lab tests and the number (and percent) of unique patients tested with the ten most common lab tests, by the BATTERY_CD field (which captures site-specific lab test names), the percentage of records this represents, and the number and percent of patients in the LAB_RESULTS table who have these results. 

```{r tab5_res}

tab5 <- odbc::dbGetQuery(con, 
  "select battery_cd,
    count(*) as nrows, 
	  count(distinct person_id) as npats
	  from lab_results
	    group by battery_cd")

tab5_sub <- tab5 %>% 
  mutate(
    pct_rec = round(nrows / tab1_lab$nrows, 4) * 100,
    pct_pat = round(npats / tab1_lab$npats, 4) * 100
  ) %>%
  arrange(desc(pct_rec)) %>% 
  head(., 10) %>% 
  select(battery_cd, nrows, pct_rec, npats, pct_pat)


```


```{r tab5_disp, results="asis"}

kable(tab5_sub, col.names = c("BATTERY_CD", "Number of Recorded Lab Tests", "Percentage of All Lab Tests", "Number of Patients Tested", "Percentage of All Patients with a Lab Test"), format.args = list(big.mark = ","))

```

##Examination of the PRESCRIBING Table (Medication Orders and Prescriptions Combined)

The following tables explore the contents of the PRESCRIBING table with a focus on trends over time and most common types of medications ordered. 

##Table 6: Annual Trends in Prescribing Data - the Total Number of Recorded Medication Orders, the Percent Change from the Previous Year, the Total Number of Patients and the Mean, Maximum and Minimum Numbers of Orders & Prescriptions per Patient

This table displays trends in PRESCRIBING record volume over time as well as patient volume represented in the table. PRESCRIBING records are attributed to a year based on RX_ORDER_DATE.


```{r tab6_res}

tab6 <- odbc::dbGetQuery(con, 
  "select year(rx_order_date) as year,
    count(*) as nrows, 
	  count(distinct person_id) as npats
	  from prescribing
	    group by year(rx_order_date)
      order by year(rx_order_date)")

tab6_avg_pats <- odbc::dbGetQuery(con, 
  "select year(rx_order_date) as year,
    count(*) as nrows
	  from prescribing
	    group by year(rx_order_date), person_id
      order by year(rx_order_date)") %>% 
  arrange(year) %>% 
  group_by(year) %>% 
  summarise_all(funs(mean, median, min, max))%>% 
  mutate(
    pats_val = paste0(median, "(", min, ", ", max, ")")
  )

tab6_all <- left_join(tab6, tab6_avg_pats, by = "year") %>% 
  mutate(
    ptc_rec = if_else(is.na(nrows / dplyr::lag(nrows)), "--", paste0(round(nrows / dplyr::lag(nrows) * 100, 0), "%"))
  ) %>% 
  select(year, nrows, ptc_rec, npats, median, min, max)

tab6_all$year <- as.character(tab6_all$year)

```

```{r tab6_disp, results="asis"}

kable(tab6_all, col.names = c("Year", "Total Number of Prescriptions", "Percent Change From Last Year", "Total Number of Patients", "Median number of Prescriptions per patient", "Minimum Number of Prescriptions Per Patient", "Maximum Number of Prescriptions Per Patient"), format.args = list(big.mark = ","))

```

##Table 7: Summary of the Ten Most Common Medications by Order Type

This table displays the top ten most common orders including count of records and patients. 

```{r tab7_res}

tab7_basis2 <- odbc::dbGetQuery(con, 
  "select top 10 GENERIC_MED_NAME,
    count(*) as nrows,
	  count(distinct person_id) as npats,
    avg(RX_DAYS_SUPPLY) as avg_days_sup,
    avg(RX_QUANTITY_NUM) as avg_rx_quant
	  from prescribing
      where RX_BASIS = '02'
	    group by GENERIC_MED_NAME
      order by nrows DESC") 

tab7_basis1 <- odbc::dbGetQuery(con, 
  "select top 10 GENERIC_MED_NAME,
    count(*) as nrows,
	  count(distinct person_id) as npats,
    avg(RX_DAYS_SUPPLY) as avg_days_sup,
    avg(RX_QUANTITY_NUM) as avg_rx_quant
	  from prescribing
      where RX_BASIS = '01'
	    group by GENERIC_MED_NAME
      order by nrows DESC") 

tab7_basis_oth <- odbc::dbGetQuery(con, 
  "select top 10 GENERIC_MED_NAME,
    count(*) as nrows,
	  count(distinct person_id) as npats,
    avg(RX_DAYS_SUPPLY) as avg_days_sup,
    avg(RX_QUANTITY_NUM) as avg_rx_quant
	  from prescribing
      where RX_BASIS in ('NI', 'UN', 'OT')
	    group by GENERIC_MED_NAME
      order by nrows DESC") 


```

```{r tab7_disp, results="asis"}
#if and else statements are used to determine if the respecitvie table has any
#rows. If not, then an error message is written indicating so.

if (nrow(tab7_basis2) != 0){
  kable(tab7_basis2, col.names = c("Medication Name (GENERIC_MED_NAME)", "Records", "Patients", "Average Days Supply", "Average Quantity"), caption = "Order to Administer (RX_BASIS = 02)", format.args = list(big.mark = ","))
} else {
  cat(noquote(paste0("Table had zero rows.  RX_BASIS = 02 likely isn't a category for this variable.", "\n")))
} 


if (nrow(tab7_basis1) != 0){
  kable(tab7_basis1, col.names = c("Medication Name (GENERIC_MED_NAME)", "Records", "Patients", "Average Days Supply", "Average Quantity"), caption = "Order to Administer (RX_BASIS = 01)", format.args = list(big.mark = ","))
} else {
  cat(noquote(paste0("Table had zero rows.  RX_BASIS = 01 likely isn't a category for this variable.", "\n")))
} 

if (nrow(tab7_basis_oth) != 0){
  kable(tab7_basis_oth, col.names = c("Medication Name (GENERIC_MED_NAME)", "Records", "Patients", "Average Days Supply", "Average Quantity"), caption = "Order to Administer (RX_BASIS = 01)", format.args = list(big.mark = ","))
} else {
   cat(noquote(paste0("Table had zero rows.  RX_BASIS = NI, UN, or OT likely aren't categories for this variable.", "\n")))
} 

```

##Examination of the PROCEDURES Table 

The following tables explore the contents of the PROCEDURES table with a focus on trends over time and most common types of procedures. 

##Table 8: Annual Trends in the Total Number of Recorded Procedures, the Mean Number of Encounters, the Mean Number of Patients and the Percent Change from the Previous Year, by Calendar Year

This table examines the volume of procedures over time and the number of encounters and patients with represented in the PROCEDURES table. Average number of procedures per encounter, and per patient, are calculated to compare procedure volume trends over time.


```{r tab8_res}

tab8 <- odbc::dbGetQuery(con, 
  "select year(procdate) as year,
    count(*) as nrows, 
    count(distinct enc_id) as nencts,
	  count(distinct person_id) as npats
	  from procedures
	    group by year(procdate)
      order by year(procdate)")

tab8_avg_encts <- odbc::dbGetQuery(con, 
  "select year(procdate) as year,
    count(*) as nrows
	  from procedures
	    group by year(procdate), enc_id
      order by year(procdate)") %>% 
  arrange(year) %>% 
  group_by(year) %>% 
  summarise_all(funs(mean, median, min, max)) %>% 
  mutate(
    enct_val = paste0(median, "(", min, ", ", max, ")")
  ) %>% 
  select(enct_val)

tab8_avg_pats <- odbc::dbGetQuery(con, 
  "select year(procdate) as year,
    count(*) as nrows
	  from procedures
	    group by year(procdate), person_id
      order by year(procdate)") %>% 
  arrange(year) %>% 
  group_by(year) %>% 
  summarise_all(funs(mean, median, min, max))%>% 
  mutate(
    pats_val = paste0(median, "(", min, ", ", max, ")")
  ) %>% 
  select(pats_val)

tab8_add <- tab8 %>% 
  bind_cols(., tab8_avg_encts, tab8_avg_pats) %>% 
  mutate(
    ptc_rec = if_else(is.na(nrows / dplyr::lag(nrows)), "--", paste0(round(nrows / dplyr::lag(nrows) * 100, 0), "%"))
  ) %>% 
  select(year, nrows, ptc_rec, nencts, npats, enct_val, pats_val)

tab8_add$year <- as.character(tab8_add$year)

```

```{r tab8_dis, results="asis"}

kable(tab8_add, col.names = c("Year", "Total Number of Recorded Procedures", "Percent Change from Previous Year", "Total Number of Encounters with Procedures", "Total Number of Patients with Procedures", "Median Number of Procedures per Encounter (Min, Max)", "Median Number of Procedures per Patient (Min, Max)"), format.args = list(big.mark = ","))

```

##Table 9:  Total Number and Percent of Recorded Encounters, Diagnoses and Procedures, by Encounter Type

This table compares the distribution of procedures, diagnoses and encounters by encounter type. Average diagnosis and procedure counts per encounter are calculated to compare diagnosis and procedure volume by encounter type.

```{r tab9_res}

tab9_enct1 <- odbc::dbGetQuery(con, 
  "select enctype,
    count(*) as nrows_enct
	  from encounters
	    group by enctype")

tab9_enct2 <- odbc::dbGetQuery(con, 
  "select count(*) as enct_tot
	  from encounters")

tab9_diag1 <- odbc::dbGetQuery(con, 
  "select enctype,
    count(*) as nrows_diag
	  from diagnoses
	    group by enctype")

tab9_diag2 <- odbc::dbGetQuery(con, 
  "select count(*) as diag_tot
	  from diagnoses")

tab9_px1 <- odbc::dbGetQuery(con, 
  "select enctype,
    count(*) as nrows_px
	  from procedures
	    group by enctype")

tab9_px2 <- odbc::dbGetQuery(con, 
  "select count(*) as px_tot
	  from procedures")

tab9_tot <- data.frame(enctype = "Total", nrows_enct = tab9_enct2$enct_tot, nrows_diag = tab9_diag2$diag_tot, nrows_px = tab9_px2$px_tot, pct_enct = NA, pct_diag = NA, pct_px = NA)

tab9_all <- tab9_enct1 %>% 
  left_join(., tab9_diag1, by = "enctype") %>% 
  left_join(., tab9_px1, by = "enctype") %>% 
  mutate(
    pct_enct = nrows_enct / tab9_enct2$enct_tot * 100,
    pct_diag = nrows_diag / tab9_diag2$diag_tot * 100,
    pct_px   = nrows_px / tab9_px2$px_tot * 100,
    enctype  = recode(enctype, "IP" = "Acute In Patient", "ED" = "Emergency Department", "AV" = "Ambulatory Visit", "TE" = "Telephone Encounter", "EM" = "E-mail Encounter", "IS" = "Non-acute Inpatient", "OE" = "Other Encounters", "LO" = "Lab Only Encounter", "RO" = "Radiology Only Encounter")
  ) %>% 
  bind_rows(., tab9_tot)  


```

```{r tab9_disp, results="asis"}

kable(tab9_all, col.names = c("Encounter Type", "Number of Encounters", "Percent of Total Encounters", "Number of Diagnoses", "Percent of Total Diagnoses", "Number of Procedures", "Percent of Total Procedures"), digits = 2, format.args = list(big.mark = ","))

```

##Table 10: Summary of the Ten Most Common Procedures - Procedure Code, Procedure Type, Total Number of Recorded Procedures and Total Number of Recorded Patients, by All or Ambulatory Encounter Type 

This table includes information on the ten most common procedures recorded in your PROCEDURE table.  Data on procedure codes, types, counts and patients are presented for all encounters and for ambulatory only encounters (ENCTYPE='AV'). 

```{r tab10_res}

tab10_pxall <- odbc::dbGetQuery(con, 
  "select px,
    px_codetype,
    count(*) as nrows_all,
    count(distinct person_id) as npats
	  from procedures
	    group by px, px_codetype") %>% 
  arrange(desc(nrows_all)) %>% 
  head(., 10)

tab10_pxav <- odbc::dbGetQuery(con, 
  "select px as px_av,
    px_codetype as px_codetype_av,
    count(*) as nrows_av,
    count(distinct person_id) as npats_av
	  from procedures
      where ENCTYPE = 'AV'
	    group by px, px_codetype") %>% 
  arrange(desc(nrows_av)) %>% 
  head(., 10)

tab10_all <- bind_cols(tab10_pxall, tab10_pxav)

```

```{r tab10_disp, results="asis"}

kable(tab10_all, col.names = c("Procedure Code", "Procedure Type", "Number of Recorded Procedures", "Number of Patients", "Procedure Code for AV", "Procedure Type for AV", "Number of Recorded Procedures for AV", "Number of Patients for AV"), format.args = list(big.mark = ","))

```

##Table 11: Summary of the Ten Most Common Procedures by Procedure Code Type, Total Number of Recorded Procedures and Total Number of Recorded Patients 

This table includes information on the ten most common procedures recorded in your PROCEDURE table.  Data on procedure codes, counts and patients are presented for CPT, HCPCS, and other code types. 

```{r tab11_res}

tab11_pxc4 <- odbc::dbGetQuery(con, 
  "select top 10 px,
    count(*) as nrows_all,
    count(distinct person_id) as npats
	  from procedures
	    where px_codetype = 'C4'
      group by px
      order by nrows_all DESC") 

tab11_pxh4 <- odbc::dbGetQuery(con, 
  "select top 10 px,
    count(*) as nrows_all,
    count(distinct person_id) as npats
	  from procedures
	    where px_codetype = 'H4'
      group by px
      order by nrows_all DESC") 

tab11_px_oth <- odbc::dbGetQuery(con, 
  "select top 10 px,
    count(*) as nrows_all,
    count(distinct person_id) as npats
	  from procedures
	    where px_codetype IN ('09', '10', '11', 'RV', 'LO', 'OT')
      group by px
      order by nrows_all DESC") 
```

```{r tab11_disp, results="asis"}
#if and else statements are used to determine if the respecitvie table has any
#rows. If not, then an error message is written indicating so.

if (nrow(tab11_pxc4) != 0){
  kable(tab11_pxc4, col.names = c("Procedure", "Records", "Patients"), caption = "Procedure Code Type = C4", format.args = list(big.mark = ","))
} else {
   cat(noquote("Table had zero rows.  PX_CODETYPE = C4 likely isn't a category for this variable.\n"))
} 


if (nrow(tab11_pxh4) != 0){
  kable(tab11_pxh4, col.names = c("Procedure", "Records", "Patients"), caption = "Procedure Code Type = H4", format.args = list(big.mark = ","))
} else {
   cat(noquote("Table had zero rows.  PX_CODETYPE = H4 likely isn't a category for this variable.\n"))
} 

if (nrow(tab11_px_oth) != 0){
  kable(tab11_px_oth, col.names = c("Procedure", "Records", "Patients"), caption = "Procedure Code Type = 09, 10, 11, RV, LO, OT", format.args = list(big.mark = ","))
} else {
   cat(noquote("Table had zero rows.  PX_CODETYPE = 09, 10, 11, RV, LO, or OT likely aren't categories for this variable.\n"))
} 


```


##Examination of the PROVIDER_SPECIALTY Table

The following tables explore the contents of the PROVIDER_SPECIALTY table with a focus on demographic data and representativeness of providers across tables.  Note that in some systems providers can be listed multiple times with multiple provider types/specialties. 

##Table 12: Number and Percent of Providers represented in ENCOUNTERS, DIAGNOSIS, LAB_RESULTS, PRESCRIBING, AND PROCEDURES Tables 

This table includes information on the number and percent of providers represented in various CHORDS VDW 3.1 tables.

```{r tab12_res}

tab12_ps <- odbc::dbGetQuery(con, 
  "select 'PROVIDER_SPECIALTY' as from_tab,
   count(distinct provider) as num_prov
     from provider_specialty")

tab12_dx <- odbc::dbGetQuery(con, 
  "select 'DIAGNOSES' as from_tab,
   count(distinct provider) as num_prov
     from diagnoses")

tab12_enct <- odbc::dbGetQuery(con, 
  "select 'ENCOUNTERS' as from_tab,
   count(distinct provider) as num_prov
     from encounters")

tab12_lab <- odbc::dbGetQuery(con, 
  "select 'LAB_RESULTS' as from_tab,
   count(distinct order_prov) as num_prov 
     from lab_results")

tab12_rx <- odbc::dbGetQuery(con, 
  "select 'PRESCRIBING' as from_tab,
   count(distinct rxmd) as num_prov 
     from PRESCRIBING")

tab12_px <- odbc::dbGetQuery(con, 
  "select 'PROCEDURES' as from_tab,
   count(distinct provider) as num_prov
     from procedures")

tab12_all <- bind_rows(tab12_ps, tab12_dx, tab12_enct, tab12_lab, tab12_rx, tab12_px) %>% 
  mutate(
    pct_prov = num_prov / tab12_ps$num_prov * 100
  )


```

```{r tab12_disp, results="asis"}

kable(tab12_all, col.names = c("Table", "Number of Providers*", "Percent of Al Providers"), digits = 2, format.args = list(big.mark = ","))

```
*Excludes Provider = 'Unknown'


##Table 13: Number and Percent of Providers by Specialty

This table includes information on the distribution of providers by specialty

```{r tab13_res}

tab13 <- odbc::dbGetQuery(con, 
  "select specialty as spec,
   count(distinct provider) as num_prov
     from provider_specialty
       group by specialty
       order by specialty") %>% 
  bind_rows(data.frame(spec = "NULL", num_prov = tab2_ps_null$SPECIALTY)) %>% 
  mutate(
    pct_prov = num_prov / tab12_ps$num_prov * 100
  ) 

tab13$spec_full <- specialties$Description[match(as.character(tab13$spec), as.character(specialties$SPECIALTY))]

tab13 <- tab13[, c(1,4,2,3)]

```
```{r tab13_disp, results = "asis"}

kable(tab13, col.names = c("Specialty Abbreviation", "Specialty", "Number of Providers", "Percent of All Providers"), digits = 2, format.args = list(big.mark = ","))

```

##Table 14: Number and Percent of Provider by Provider Type 

This table includes information on the distribution of providers by type.


```{r tab14_res}

tab14 <- odbc::dbGetQuery(con, 
  "select provider_type as prov_type,
   count(distinct provider) as num_prov
     from provider_specialty
       group by provider_type
       order by provider_type") %>% 
  bind_rows(data.frame(prov_type = "NULL", num_prov = tab2_ps_null$PROVIDER_TYPE)) %>% 
  mutate(
    pct_prov = num_prov / tab12_ps$num_prov * 100
  ) 

tab14$spec_full <- prov_type$V2[match(as.character(tab14$prov_type), as.character(prov_type$V1))]

tab14 <- tab14[, c(1,4,2,3)]

```

```{r tab14_disp, results = "asis"}

kable(tab14, col.names = c("Provider Abbreviation", "Provider Type", "Number of Providers", "Percent of All Providers"), digits = 2, format.args = list(big.mark = ","))

```

##Table 15: Annual Trends in Self-Reported Tobacco Use - the Distribution of Responses by Tobacco Behavior Categories, by Calendar Year	

This table includes information on the distribution of SOCIAL_HISTORY records by TOBACCO_USER value.

```{r tab15_res}

tab15 <- odbc::dbGetQuery(con, 
  "select year(contact_date) as year,
   count(*) as num_rec,
   SUM(CASE WHEN TOBACCO_USER = 'Y' THEN 1 ELSE 0 END) as smoker,
   SUM(CASE WHEN TOBACCO_USER = 'I' THEN 1 ELSE 0 END) as infreq,
   SUM(CASE WHEN TOBACCO_USER = 'Q' THEN 1 ELSE 0 END) as former,
   SUM(CASE WHEN TOBACCO_USER = 'P' THEN 1 ELSE 0 END) as passive,
   SUM(CASE WHEN TOBACCO_USER = 'N' THEN 1 ELSE 0 END) as never,
   SUM(CASE WHEN TOBACCO_USER IN ('X', 'U') THEN 1 ELSE 0 END) as unknown,
   SUM(CASE WHEN TOBACCO_USER IS NULL THEN 1 ELSE 0 END) as t_null
     from social_history
       group by year(contact_date)
       order by year(contact_date)")

tab15$year <- as.character(tab15$year)

```

```{r tab15_disp, results="asis"}
kable(tab15, col.names = c("Year", "Records", "Yes/Current", "Infrequent", "Quit/Former", "Passive/Second Hand Smoke Exposure", "Never", "Not Asked/Unknown", "Null"), format.args = list(big.mark = ","))

```

##Table 16: Annual Trends in Self-Reported Alcohol Use - the Distribution of Responses by Alcohol Behavior Categories, by Calendar Year	

This table includes information on the distribution of SOCIAL_HISTORY records by ALCOHOL_USER value.


```{r tab16_res}

tab16 <- odbc::dbGetQuery(con, 
  "select year(contact_date) as year,
   count(*) as num_rec,
   SUM(CASE WHEN ALCOHOL_USER = 'Y' THEN 1 ELSE 0 END) as alc_use,
   SUM(CASE WHEN ALCOHOL_USER = 'Q' THEN 1 ELSE 0 END) as former,
   SUM(CASE WHEN ALCOHOL_USER = 'N' THEN 1 ELSE 0 END) as never,
   SUM(CASE WHEN ALCOHOL_USER IN ('X', 'U') THEN 1 ELSE 0 END) as unknown,
   SUM(CASE WHEN ALCOHOL_USER IS NULL THEN 1 ELSE 0 END) as t_null
     from social_history
       group by year(contact_date)
       order by year(contact_date)")

tab16$year <- as.character(tab16$year)

odbc::dbGetQuery(con,
  "select top 10 *
from social_history")

```

```{r tab16_disp, results="asis"}
kable(tab16, col.names = c("Year", "Records", "Yes/Current", "Quit/Former",  "Never", "Not Asked/Unknown", "Null"), format.args = list(big.mark = ","))

```

##Table 17: Annual Trends in Self-Reported Illicit Drug Use - the Distribution of Responses by Illicit Drug Use Behavior Categories, by Calendar Year	

This table includes information on the distribution of SOCIAL_HISTORY records by ILL_DRUG_USER value.

```{r tab17_res}

tab17 <- odbc::dbGetQuery(con, 
  "select year(contact_date) as year,
   count(*) as num_rec,
   SUM(CASE WHEN ILL_DRUG_USER = 'Y' THEN 1 ELSE 0 END) as ill_drug,
   SUM(CASE WHEN ILL_DRUG_USER = 'Q' THEN 1 ELSE 0 END) as former,
   SUM(CASE WHEN ILL_DRUG_USER = 'N' THEN 1 ELSE 0 END) as never,
   SUM(CASE WHEN ILL_DRUG_USER IN ('X', 'U') THEN 1 ELSE 0 END) as unknown,
   SUM(CASE WHEN ILL_DRUG_USER IS NULL THEN 1 ELSE 0 END) as t_null
     from social_history
       group by year(contact_date)
       order by year(contact_date)")

tab17$year <- as.character(tab17$year)

```

```{r tab17_disp, results="asis"}
kable(tab17, col.names = c("Year", "Records", "Yes/Current", "Quit/Former",  "Never", "Not Asked/Unknown", "Null"), format.args = list(big.mark = ","))

```

####Total program run time:
```{r calc_runtime}
# close odbc
DBI::dbDisconnect(con)

endTime <- Sys.time()
runtime <- endTime - startTime 
```


Query run time = `r runtime` minutes
